% appA-prelims.tex -- Additional preliminaries.
\section{Additional preliminaries}\label{app:prelims}

\paragraph{Negative association: the closure properties we use.}
We use two standard closure properties of negatively associated families (\Cref{def:NA}), both from the original treatment of
\cite{JoagDevProschan1983}, together with the one tail inequality they imply. Recall that a random vector $Z=(Z_1,\dots,Z_m)$
is NA if, for every pair of disjoint index sets $I,J\subseteq[m]$ and every pair $f,g$ of coordinatewise-nondecreasing
functions of the corresponding subvectors $Z_I,Z_J$,
\[
  \E\big[f(Z_I)\,g(Z_J)\big]\ \le\ \E\big[f(Z_I)\big]\,\E\big[g(Z_J)\big],
\]
so the two resulting functions are nonpositively correlated. We record four facts. The first two are background that pins down
the meaning, and the second two are the closure steps the row-load argument actually invokes.

The base case (P1, independence implies NA) is that a vector of mutually independent coordinates is NA: for independent
$Z_I,Z_J$ on disjoint blocks the covariance in the display is exactly zero, so the inequality holds with equality. This is the
seed from which the structured NA vectors below are built.

The factorization fact (P4, upper-tail MGF) is the single analytic consequence we use downstream and is stated separately at
the end of this paragraph; it follows from the defining inequality applied inductively to the nondecreasing functions
$e^{tZ_i}$.

The first closure step we apply (property P7 of \cite{JoagDevProschan1983}, here (P2), closure under independent union) says
that NA is preserved by concatenating independent NA vectors: if $X$ and $Y$ are each NA and are independent of each other,
then the stacked vector $(X,Y)$ is NA. The reason is that any two monotone functions on disjoint index sets either read
coordinates of a single block, where the NA of that block applies, or split across the two blocks, where independence makes
the covariance vanish; in every case the defining inequality holds. Combining (P1) and (P2) one gets that a concatenation of
independent NA blocks is NA, which is how a product structure enters.

The second closure step (property P6 of \cite{JoagDevProschan1983}, here (P3), monotone functions of disjoint blocks) says
that NA is preserved under taking nondecreasing functions of disjoint groups of coordinates: if $(Z_1,\dots,Z_m)$ is NA and
$g_1,\dots,g_r$ are coordinatewise-nondecreasing functions, each depending only on its own \emph{disjoint} block of indices,
then $(g_1(Z),\dots,g_r(Z))$ is again NA. This is immediate from the definition, since any monotone function of a
subcollection of the $g_i$ is itself a monotone function of a union of the underlying disjoint blocks, and that union is
disjoint from the corresponding union for the complementary subcollection, so the defining inequality for the $g_i$ is a
special case of the defining inequality for $Z$.

These two facts are exactly what is needed to certify that the row loads of a random column-sampled matrix are NA. Fix a
subset $S$ of $s$ columns. The columns are sampled independently, and within a single column the entries form an NA vector
(a vector of indicators of a uniform $d$-subset is NA by the permutation-distribution case of \cite{JoagDevProschan1983}).
Applying (P7) across the $s$ independent columns of $S$ therefore yields an NA vector on the product index set
$[n]\times S$. Now apply (P6) with the functions $g_i(\cdot)=\sum_{t\in S}A_{it}$, one per row $i$. The function $g_i$ reads
only the block $\{(i,t):t\in S\}$, these blocks are disjoint across the rows $i$, and each $g_i$ is nondecreasing in its
entries, so the hypotheses of (P6) hold and the row-load vector $\{r_i(S)\}_{i\in[n]}$ is NA.

The single inequality we actually invoke downstream is the upper-tail (MGF) direction of NA, fact (P4): for nondecreasing
functions $f_1,\dots,f_n$ of the NA coordinates $Y_1,\dots,Y_n$ and any $t>0$,
\[
  \E\Big[\textstyle\prod_i f_i(Y_i)\Big]\ \le\ \prod_i\E[f_i(Y_i)],
\]
which is the statement that the moment generating function of a sum of NA coordinates factorizes as an upper bound. To see how
it follows from the definition, peel off one factor at a time: write $\prod_{i\le n}f_i(Y_i)=f_n(Y_n)\cdot\prod_{i<n}f_i(Y_i)$
and note that $\prod_{i<n}f_i(Y_i)$ is a nondecreasing function of the block $\{Y_i:i<n\}$, while $f_n(Y_n)$ is a
nondecreasing function of the disjoint singleton $\{Y_n\}$. The defining inequality applied to this split gives
$\E[\prod_{i\le n}f_i(Y_i)]\le\E[\prod_{i<n}f_i(Y_i)]\,\E[f_n(Y_n)]$, and induction on $n$ closes the product. Taking
$f_i(\cdot)=e^{t\,(\cdot)}$, which is nondecreasing for $t>0$, yields the factorization $\E[e^{t\sum_iY_i}]\le\prod_i\E[e^{t
Y_i}]$. It is the direction of NA that makes Chernoff-type bounds go through as if the coordinates were independent. We must
check that we apply it with the inequality pointing the right way, that is, with genuinely nondecreasing $f_i$. In both estimates of
\Cref{app:structural-full} the summand $Y_i$ is a nondecreasing function of the load $r_i$: for the dense estimate it is
$\exp(-c/r)\1[r>0]$, which increases as $r$ increases (a larger load makes the exponential closer to $1$ and keeps the
indicator on); for the sparse estimate it is the threshold indicator $\1[r>U]$, which is also nondecreasing in $r$. Hence in
each case $e^{tY_i}$ is a nondecreasing function of $Y_i$ and so of $r_i$, the factorization above applies in the correct
(upper-bound) direction, and the moment generating function of $\sum_i Y_i$ is dominated by the product of the per-coordinate
moment generating functions.

\paragraph{The exact-equality form of Lovett--Meka.}
In \Cref{lem:LM} a threshold $c_j=0$ on a vector $v_j$ contributes $e^{-0}=1$ to the budget $\sum_j e^{-c_j^2/16}\le m/16$ and
enforces $\langle v_j,x-x_0\rangle=0$ exactly. We use this only for $v_j=\mathbf 1$ (the all-ones vector), so the budget
spent is $1$; together with the row constraints ($\le m/64$) it stays under $m/16$ as soon as $m\ge 22$. The partial-coloring
walk is run as a Brownian motion constrained to the subspace $\{x:\langle\mathbf 1,x\rangle=\langle\mathbf 1,x_0\rangle\}$,
the standard way to impose an exact linear equality; the cardinality of the output is therefore exact up to the $\Oh(1)$
coordinates left fractional when the alive set drops below $22$, which are rounded by hand at additive cost $\Oh(1)$.

\paragraph{Poisson tail, full statement.}
We record the upper tail of a Poisson variable in the Bennett form used in \eqref{eq:poisson-tail}, together with its
derivation, its two regimes, and the transfer to the binomial. Let $X\sim\Pois(\mu)$ and let $t\ge0$. For any $\theta>0$ the
Chernoff bound gives
\[
  \Pr(X\ge\mu+t)\ \le\ e^{-\theta(\mu+t)}\,\E[e^{\theta X}]\ =\ \exp\!\big(\mu(e^{\theta}-1)-\theta(\mu+t)\big),
\]
using the Poisson moment generating function $\E[e^{\theta X}]=\exp(\mu(e^{\theta}-1))$. Minimizing the exponent over
$\theta$ sets $e^{\theta}=1+t/\mu$, that is $\theta=\log(1+t/\mu)$, and substituting this optimal $\theta$ collapses the
exponent to $-\mu\,h(t/\mu)$ with
\[
  h(u)\ =\ (1+u)\log(1+u)-u\ \ge\ 0 .
\]
To see the collapse of the exponent in detail, substitute $e^{\theta}=1+t/\mu$ into $\mu(e^{\theta}-1)-\theta(\mu+t)$. The
first term becomes $\mu(t/\mu)=t$, and the second becomes $\log(1+t/\mu)\,(\mu+t)$, so the exponent equals
\[
  t-(\mu+t)\log\!\Big(1+\frac t\mu\Big)\ =\ -\mu\Big[\big(1+\tfrac t\mu\big)\log\big(1+\tfrac t\mu\big)-\tfrac t\mu\Big]\ =\ -\mu\,h(t/\mu),
\]
which is the rate function written out: $\mu\,h(t/\mu)=(\mu+t)\log(1+t/\mu)-t$.
Thus $\Pr(X\ge\mu+t)\le\exp(-\mu\,h(t/\mu))$, which is the upper bound quoted in \eqref{eq:poisson-tail}. The function $h$ is
nonnegative because $h(0)=0$ and $h'(u)=\log(1+u)\ge0$, so it is increasing on $[0,\infty)$. A matching lower bound of the
same exponential order holds for $\mu\ge1$ and $t\le\mu^{2}$ by the local central limit theorem, which gives the Poisson
point mass at $\mu+t$ up to polynomial factors and lets one sum the tail with no loss in the exponent; this is what makes the
estimate tight and not merely an upper bound.

The two regimes of \eqref{eq:poisson-tail} are read off from the behavior of $h$. For small deviations, $u\le1$, a Taylor
expansion of $h(u)=(1+u)\log(1+u)-u$ around $u=0$, using $\log(1+u)=u-\tfrac12u^2+\tfrac13u^3-\cdots$, gives
$h(u)=\tfrac12 u^2-\tfrac16u^3+\cdots\asymp u^2$, so with $u=t/\mu$ the rate is $\mu\,h(t/\mu)\asymp\mu\cdot(t/\mu)^2=t^2/\mu$
and the bound is the Gaussian one $\exp(-\Th(t^2/\mu))$. For large deviations, $u\ge1$, the leading term is $h(u)\asymp
u\log(1+u)$, so $\mu\,h(t/\mu)\asymp\mu\cdot(t/\mu)\log(1+t/\mu)=t\log(1+t/\mu)$ and the bound becomes the Poissonian
$\exp(-\Th(t\log(1+t/\mu)))$, which decays faster than any Gaussian. The two expressions $t^2/\mu$ and $t\log(1+t/\mu)$ agree
in order at $u\asymp1$, since there $\log(1+u)\asymp u$ makes $t\log(1+t/\mu)\asymp t\cdot(t/\mu)=t^2/\mu$. This crossover at
$u\asymp1$, that is at $t\asymp\mu$, is exactly the point where the deviation stops being typical-scale and the heavier
Poisson tail takes over.

The same estimate transfers verbatim to the binomial, by the same Chernoff optimization. For $X\sim\Bin(m,p)$ with mean
$\mu=mp$, the moment generating function is $\E[e^{\theta X}]=(1-p+pe^{\theta})^m$, and $\log\E[e^{\theta X}]=m\log(1-p+pe^
{\theta})\le m\,p(e^{\theta}-1)=\mu(e^{\theta}-1)$, using $\log(1+x)\le x$ with $x=p(e^{\theta}-1)$. Thus the binomial
cumulant generating function is dominated by the Poisson one of the same mean, the optimization over $\theta$ goes through
unchanged, and Bennett's inequality gives the identical $\exp(-\mu\,h(t/\mu))$ upper tail. This is the binomial-to-Poisson
transfer: the binomial tail is never heavier than the Poisson tail of the matching mean. The matching lower bound of the same
order holds once the variance is at least constant, $mp(1-p)\ge1$, so that the local central limit theorem applies and
supplies the Poisson point mass at $\mu+t$ up to polynomial factors. This variance condition holds throughout our
application, because the relevant binomial means are $\Th(\lam)=\Om(1)$ in the regime $k\le d$, and the success probabilities
are bounded away from $1$, so $mp(1-p)=\Th(\lam)\ge1$.
