% appB-offline.tex -- full deferred proofs of Section 3.
\section{Full proof of the offline characterization}\label{app:offline}

Throughout, $A\in\{0,1\}^{n\times T}$ is the incidence matrix ($A_{it}=\1[i\in\supp(v_t)]$), $r_i(S)=\sum_{t\in S}A_{it}$,
$\mu_s=ds/n$, $\ell_s=\log(eT/s)$. We assume $\tau_-\,n\le T\le\tau_+\,n$ for absolute constants $0<\tau_-\le\tau_+$, so
$\ell_s=\Th(\log(n/s))$ and $\ell_s\ge\log e=1$.

\subsection{The lower bound (\Cref{lem:offline-lower})}\label{app:offline-lower-full}

\paragraph{Poisson tails and the inversion.}
For $X\sim\Pois(\mu)$, the Cram\'er rate at $\mu+t$ is $I(t)=\mu\,h(t/\mu)$ with $h(u)=(1+u)\log(1+u)-u$. Since
$h(u)=\Th(u^2)$ for $u\le 1$ and $h(u)=\Th(u\log(1+u))$ for $u\ge 1$, we get \eqref{eq:poisson-tail}. Solving rate $=\Lscale$:
in the regime $t\lesssim\mu$, $t^2/\mu\asymp\Lscale$ gives $t\asymp\sqrt{\mu\Lscale}$; in the regime $t\gtrsim\mu$, setting
$t=c\,\Lscale/\log(e+\Lscale/\mu)$ and $z=\Lscale/\mu\ge 1$ gives $t\log(1+t/\mu)\asymp \mu\cdot\frac{z}{\log(e+z)}\cdot\log\!\big(1+\tfrac{z}{\log(e+z)}\big)\asymp\mu z=\Lscale$, because $\log(1+z/\log(e+z))\asymp\log(e+z)$ for $z\ge 1$. Hence the deviation at which a single $\Pois(\mu)$ load reaches rate $\Lscale$ is $t^*(\mu,\Lscale)=\Th(\sqrt{\mu\Lscale}+\Lscale/\log(e+\Lscale/\mu))=\Th(\Psifn(\mu,\Lscale))$.

\paragraph{Step 1: means.}
Fix a coloring with means $\mu_c=m_c d/n$, $\sum_c\mu_c=k\lam$. The loads are independent with $\E Y_c=\mu_c$ and
$\Var Y_c=\mu_c(1-d/n)=\Th(\mu_c)$. We do not assume balance; instead we lower-bound the range by a dichotomy on the means.
Throughout, $t^*=t^*(\lam,\log k)=\Th(\Psifn(\lam,\log k))$ is the deviation of the Poisson inversion above, and
$t^*\ge 4\sqrt{\lam\log k}$ (since $\Psifn\ge\sqrt{\lam\log k}$ and $\log k$ is large).

\paragraph{Step 2: Poissonization.}
Let $B$ be the $n\times T$ matrix with i.i.d.\ $\Bin(1,p)$ entries, $p=d/n$. Conditioning each column of $B$ on having
exactly $d$ ones yields $\mu_{d,n}$, and by Stirling $\Pr(\Bin(n,p)=d)\ge c_0/\sqrt{d+1}$, so $\Pr(\mathcal C)\ge(c_0/\sqrt{d+1})^T$
where $\mathcal C$ is the event that all columns have weight $d$. Hence for any event $E$,
$\Pr_{\mathrm{exact}}(E)\le\Pr_B(E)/\Pr_B(\mathcal C)\le(C\sqrt{d+1})^T\Pr_B(E)$. Under $B$ the $k$ loads at a fixed
coordinate, $Y_c=\sum_{t\in\chi^{-1}(c)}B_{it}\sim\Bin(m_c,p)$, are independent, with $\E Y_c=\mu_c=m_c p$ and
$\Var Y_c=\Th(\mu_c)$ and $\sum_c\mu_c=k\lam$.

\paragraph{Step 3: per-coordinate spread.}
Fix a coordinate. The $k$ loads $Y_c\sim\Bin(m_c,p)$ are independent with means $\mu_c$ summing to $k\lam$. Write
$t^*=\Psifn(\lam,\log k)$. We prove the following lower tail, under the standing hypothesis $\lam\ge 1$.

\begin{remark}[Necessity of $\lam\ge1$]\label{rem:lam-ge-1}
Some lower bound on $\lam$ is unavoidable. For every profile all $k$ loads vanish together with probability
$\prod_c(1-p)^{m_c}=(1-p)^T=e^{-(1+o(1))k\lam}$, an event of range $0$, so $\Pr(\rangeop_i\le\tfrac14 t^*)\ge e^{-(1+o(1))k\lam}$.
Since $k\lam=Tp=\Th(d)$, no bound $e^{-k^{\Om(1)}}$ can hold once $\lam=o(1)$ (equivalently $k=\omega(d)$). We therefore
assume $\lam\ge1$, which in the range $k\le d$ holds whenever $T\ge n$; it excludes only the top sliver $k=\omega(Td/n)$ of
the palette.
\end{remark}

\begin{lemma}[Occupancy spread, lower tail]\label{lem:occ-spread}
Assume $\lam\ge1$ and $p(k\lam+\log^2 k)=o(1)$. Let $Y_1,\dots,Y_k$ be independent with $Y_c\sim\Bin(m_c,p)$, $\mu_c=m_c p\ge0$,
$\sum_c\mu_c=k\lam$. Then $\Pr\big(\max_c Y_c-\min_c Y_c\le\tfrac14 t^*\big)\le e^{-c k^{1/4}}$ for an absolute $c>0$; in
particular the rate is at least $k^{1/5}$.
\end{lemma}

The engine is a spread bound for bins of comparable mean, which we prove from a reverse Cram\'er estimate. Write
$J_\mu(x)=x\log(x/\mu)-x+\mu$ for the Poisson rate.

\begin{lemma}[Reverse Cram\'er]\label{lem:reverse-cramer}
Let $X\sim\Bin(N,p)$ with $\mu=Np$, and let $x$ be a nonnegative integer with $p(\mu+x^2/\mu)=o(1)$. Put $q=x/N$,
$t=\log\frac{q(1-p)}{p(1-q)}$, $\sigma^2=Nq(1-q)$. Then $\Pr(X\ge x)\ge e^{-J_\mu(x)-\Oh(\log(1+|t|\sigma))}$ for $x\ge\mu$,
and $\Pr(X\le x)\ge e^{-J_\mu(x)-\Oh(\log(1+|t|\sigma))}$ for $x\le\mu$.
\end{lemma}

\begin{proof}
Tilt $\Bin(N,p)$ to $Q=\Bin(N,q)$, under which $x$ is the mean. The likelihood ratio is
$\frac{dP_p}{dQ}(z)=\exp(-N D(q\Vert p)-t(z-x))$ with $D$ the binary relative entropy. A Stirling estimate gives $Q$ at least
$\tfrac{c}{1+|t|\sigma}$ mass on an adjacent one-sided integer interval where $|t(z-x)|\le1$, so the corresponding $P_p$-mass
is at least $\tfrac{c}{1+|t|\sigma}e^{-ND(q\Vert p)}$. Finally $ND(q\Vert p)=J_\mu(x)+\Oh(p(\mu+x^2/\mu))=J_\mu(x)+o(1)$.
\end{proof}

\begin{lemma}[Comparable-mean spread]\label{lem:comparable}
Let $m\to\infty$, $\ell=\log m$, and let $Z_1,\dots,Z_m$ be independent with $Z_j\sim\Bin(N_j,p)$, means $\mu_j\in[M,2M]$,
$M\ge\tfrac12$, and $p(M+\ell^2)=o(1)$. Then $\Pr\big(\max_j Z_j-\min_j Z_j\le\tfrac25\Psifn(M,\ell)\big)\le e^{-c\sqrt m}$ for
an absolute $c>0$.
\end{lemma}

\begin{proof}
Let $A,B$ index the $h=\lfloor m/2\rfloor$ smallest and largest means; put $a=\max_{j\in A}\mu_j\le b=\min_{j\in B}\mu_j$, so
$M\le a\le b\le 2M$, and write $\psi=\Psifn(M,\ell)$. With $g(u)=\sqrt u+1/\log(e+1/u)$ (so $\psi/\ell=g(M/\ell)+o(1)$),
one-variable calculus gives
\begin{equation}\label{eq:rate-small}
  J_{u\ell}\big(\tfrac25\ell g(u)\big)\le\tfrac25\ell\quad\text{whenever }\tfrac25 g(u)\ge u,
\end{equation}
\begin{equation}\label{eq:rate-large}
  J_{u\ell}\big(u\ell\pm\tfrac15\ell g(u)\big)\le\tfrac18\ell\quad\text{for }u\ge\tfrac{9}{20}.
\end{equation}

\emph{Case $a\le\tfrac{9}{20}\ell$.} Each $j\in A$ has $\Pr(Z_j=0)=(1-p)^{N_j}=e^{-(1+o(1))\mu_j}\ge m^{-9/20-o(1)}$, so
$\Pr(\min_{j\in A}Z_j>0)\le e^{-m^{11/20-o(1)}}$. Let $r=\lceil\tfrac25\psi\rceil$. By \Cref{lem:reverse-cramer} and
\eqref{eq:rate-small}, each $j\in B$ has $\Pr(Z_j\ge r)\ge m^{-2/5-o(1)}$, so $\Pr(\max_{j\in B}Z_j<r)\le e^{-m^{3/5-o(1)}}$.
Off these events the minimum is $0$ and the maximum is $\ge r$, giving spread $\ge\tfrac25\psi$. When $\psi>M$ the minimum
is simply $0$ and is never asked to go negative.

\emph{Case $a>\tfrac{9}{20}\ell$.} Put $s=\tfrac15\psi\le\tfrac15\Psifn(a,\ell)$. By \Cref{lem:reverse-cramer} and
\eqref{eq:rate-large}, each $j\in A$ has $\Pr(Z_j\le a-s)\ge m^{-1/8-o(1)}$, and each $j\in B$ has $\Pr(Z_j\ge b+s)\ge
m^{-1/8-o(1)}$ (the displacement-$s$ rate at mean $b\ge a$ is no larger). By independence,
$\Pr(\min_{j\in A}Z_j>a-s)+\Pr(\max_{j\in B}Z_j<b+s)\le 2e^{-m^{7/8-o(1)}}$. Off these events the spread is
$\ge(b+s)-(a-s)\ge 2s=\tfrac25\psi$.
\end{proof}

\begin{proof}[Proof of \Cref{lem:occ-spread}]
Two cases on the largest mean $\mu_h:=\max_c\mu_c$.

\emph{Heavy bin: $\mu_h\ge k^{1/4}\lam$.} Let $S=\{c:\mu_c\le\lam\}$. From $\sum_c(\mu_c-\lam)=0$,
$|S|\lam\ge\sum_{c\in S}(\lam-\mu_c)=\sum_{c\notin S}(\mu_c-\lam)\ge\mu_h-\lam$, so $|S|\ge k^{1/4}-1$. A Chernoff bound gives
$\Pr(Y_h<\mu_h/2)\le e^{-\mu_h/8}\le e^{-k^{1/4}/8}$ (using $\lam\ge1$), and each $c\in S$ has $\Pr(Y_c>2\lam)\le\tfrac12$
(Markov), so $\Pr(\min_{c\in S}Y_c>2\lam)\le 2^{-|S|}\le e^{-ck^{1/4}}$. Off these events $\rangeop_i\ge\tfrac12 k^{1/4}\lam-2\lam$,
which exceeds $\tfrac14 t^*$ because $t^*/\lam\le 1+\sqrt{\log k}+\log k=\Oh(\log k)$ for $\lam\ge1$ and $k^{1/4}=\omega(\log k)$.

\emph{No heavy bin: $\mu_h<k^{1/4}\lam$.} Let $H=\{c:\mu_c\ge\tfrac{9}{10}\lam\}$; the discarded bins carry mean
$<\tfrac{9}{10}k\lam$, so $\sum_{c\in H}\mu_c>\tfrac1{10}k\lam$ and $|H|\ge\tfrac1{10}k^{3/4}$. Split
$[\tfrac{9}{10}\lam,k^{1/4}\lam)$ into $\Oh(\log k)$ dyadic bands; the densest band $B$ has $m:=|B|\ge ck^{3/4}/\log k=k^{3/4-o(1)}$
bins of comparable mean $M\in[\tfrac{9}{10}\lam,k^{1/4}\lam)$, with $\ell=\log m\ge\tfrac{7}{10}\log k$. A direct estimate gives
$\Psifn(M,\ell)\ge\tfrac58\,t^*$: the constant and Gaussian terms only grow, and for the Bennett term
$\log(e+\tfrac{10}{9}\tfrac{\log k}{\lam})\le(1+\log\tfrac{10}{9})\log(e+\tfrac{\log k}{\lam})$ with
$\tfrac{7}{10}/(1+\log\tfrac{10}{9})>\tfrac58$. By \Cref{lem:comparable} (whose hypothesis $p(M+\ell^2)=o(1)$ is the assumed
$p(k\lam+\log^2 k)=o(1)$ since $M\le k\lam$ and $\ell\le\log k$),
$\Pr(\max_{c\in B}Y_c-\min_{c\in B}Y_c\le\tfrac25\Psifn(M,\ell))\le e^{-c\sqrt m}$, and
$\tfrac25\Psifn(M,\ell)\ge\tfrac25\cdot\tfrac58\,t^*=\tfrac14 t^*$. The full range dominates the range over $B$, so
$\Pr(\rangeop_i\le\tfrac14 t^*)\le e^{-c\sqrt m}=e^{-k^{3/8-o(1)}}$.

Both cases give $\Pr(\rangeop_i\le\tfrac14 t^*)\le e^{-ck^{1/4}}$, and $k^{1/4}\ge k^{1/5}$.
\end{proof}

In our model the two hypotheses of \Cref{lem:occ-spread} hold: $\lam=Td/(nk)\ge1$ in the stated range (\Cref{rem:lam-ge-1}),
and $p(k\lam+\log^2 k)=\tfrac dn(\Th(d)+\Oh(\log^2 d))=\Oh(d^2/n)=o(1)$ since $d\le\Hscale^2/\log\Hscale$. Step~2 placed us in
the Bernoulli model $B$, whose rows are independent; the events $\{\rangeop_i\le D\}_{i\in[n]}$ depend on disjoint rows of
$B$, so they are independent and $\Pr_B(\max_i\rangeop_i\le\tfrac14 t^*)\le e^{-qn}$ by \Cref{lem:occ-spread}. Set
$D=\tfrac14 t^*=\Th(\Psifn)$.

\paragraph{Step 4: union over colorings.}
Step 3 gives a per-coordinate rate $q\ge k^{1/5}$. Combining Steps 2--3 and union-bounding over the at most $k^T$ colorings,
\[
  \Pr_{\mathrm{exact}}\big(\exists\chi:\ \max_i\rangeop_i(T)\le D\big)\ \le\ (C\sqrt{d+1})^T\,k^T\,e^{-q n}
  \ =\ \exp\!\big(\Oh(n\log(dk))-\Om(n\,k^{1/5})\big).
\]
By the hypothesis $k\ge C_0\log^5(dk)$, $k^{1/5}\ge C_0^{1/5}\log(dk)$ dominates the $\Oh(\log(dk))$ per-coordinate cost, so
the probability is $\exp(-\Om(n\,k^{1/5}))=o(1)$. Therefore $\offdisc_k>D=\tfrac14 t^*\ge c\,\Psifn(\lam,\log k)$ \whp. The three
regimes \eqref{eq:Psi-regimes} are the three behaviours of $t^*(\lam,\log k)$: Gaussian $\sqrt{\lam\log k}$ for
$\lam\gtrsim\log k$, the Bennett term $\log k/\log((\log k)/\lam)$ for $1\ll\lam\ll\log k$, and $\log k/\loglog k$ for
$\lam=\Th(1)$. \qed

\subsection{The structural lemma (\Cref{lem:structural})}\label{app:structural-full}

The structural lemma asserts that, simultaneously for every subset $S$ of columns, the row loads $r_i(S)$ are not too
concentrated on any small set of rows. The proof has the shape one expects for a statement of this kind. We first show that
a single row is unlikely to be heavily loaded, controlling the expected number of overloaded rows by a moment computation;
we then upgrade this expectation bound to a high-probability statement using a tail inequality that survives the dependence
between rows; finally we pay for the union over all $\binom{T}{s}$ choices of $S$ and over all $s$. The one structural
ingredient that makes the upgrade possible is negative association, which lets us treat the row loads as if they were
independent for the purpose of upper-tail bounds.

\paragraph{Negative association.}
We begin by recording why the row loads are negatively associated, since this is what licenses the Chernoff-type
factorization used below. The indicator vector of a uniform $d$-subset of $[n]$ is strongly Rayleigh, and strong Rayleigh
measures are negatively associated across coordinates \cite{BorceaBrandenLiggett2009}. The $s$ columns indexed by $S$ are
drawn independently, so concatenating their indicator vectors gives a negatively associated vector on the product index set
$\{(i,t):i\in[n],t\in S\}$; independence of the blocks preserves NA. Now each row load $r_i(S)=\sum_{t\in S}A_{it}$ is a
nondecreasing function of the block $\{(i,t):t\in S\}$, and distinct rows $i$ depend on disjoint blocks. Negative
association is closed under taking nondecreasing functions of disjoint blocks of variables, so the family
$\{r_i(S)\}_{i\in[n]}$ is itself NA (\Cref{app:prelims}). This is exactly the property we need: for nondecreasing functions
of the $r_i$, the upper tail of a sum factorizes as if the rows were independent.

\paragraph{The two tail estimates.}
Fix $S$ and write $s=|S|$. Each row load $r_i(S)$ is a sum of $s$ independent $\Bin(1,d/n)$ indicators, so
$r_i(S)\sim\Bin(s,d/n)$ with mean $\mu_s=ds/n$. Fix a large constant $B$; it will be tuned at the very end so that the union
bound closes. The level above which a row counts as overloaded changes character depending on whether the mean $\mu_s$ is
above or below the logarithmic budget $\ell_s$, so we treat the two regimes separately. In each case we exhibit a
nonnegative variable $Y_i$, a nondecreasing function of $r_i$, whose expectation is at most $e^{-(B-1)\ell_s}$; this is the
per-row input the high-probability step consumes.

\emph{Dense case $\mu_s\ge\ell_s$.} Here the natural overload scale is the multiplicative one, so we use a soft exponential
test variable in place of a hard threshold. Let $Y_i=\exp(-K^2\mu_s\ell_s/(16\,r_i))\1[r_i>0]$, which is increasing in
$r_i$ and bounded by $1$. Split on whether $r_i$ exceeds $b\mu_s$ for a constant $b$ to be chosen. On the event
$\{0<r_i\le b\mu_s\}$, since $r_i\le b\mu_s$ enlarges the denominator and so shrinks the magnitude of the negative exponent,
the exponent $-K^2\mu_s\ell_s/(16 r_i)$ is at most
\[
  -\frac{K^2\mu_s\ell_s}{16\,r_i}\ \le\ -\frac{K^2\mu_s\ell_s}{16\,b\mu_s}\ =\ -\frac{K^2\ell_s}{16b},
\]
so $Y_i\le e^{-K^2\ell_s/(16 b)}$ there. On the complementary event $\{r_i>b\mu_s\}$ we only use $Y_i\le 1$, but this event
is itself rare: writing $r_i\sim\Bin(s,d/n)$ with mean $\mu_s$, the multiplicative Chernoff bound
$\Pr(r_i>(1+\delta)\mu_s)\le\exp(-\mu_s\,((1+\delta)\log(1+\delta)-\delta))$ at $1+\delta=b$ gives
\[
  \Pr(r_i>b\mu_s)\ \le\ \exp\!\big(-\mu_s\,(b\log b-b+1)\big)\ =\ e^{-c_b\mu_s}\ \le\ e^{-c_b\ell_s},
  \qquad c_b:=b\log b-b+1,
\]
the last step using $\mu_s\ge\ell_s$ and $c_b>0$ for $b>1$. Combining the two events,
\[
  \E Y_i\ \le\ e^{-K^2\ell_s/(16b)}\cdot\Pr(0<r_i\le b\mu_s)+1\cdot\Pr(r_i>b\mu_s)\ \le\ e^{-K^2\ell_s/(16b)}+e^{-c_b\ell_s}.
\]
We now fix constants in order: choose $b$ large enough that $c_b=b\log b-b+1\ge B$, and then choose $K$ large enough that
$K^2/(16b)\ge B$. Both summands are then at most $e^{-B\ell_s}$, so $\E Y_i\le 2e^{-B\ell_s}\le e^{-(B-1)\ell_s}$, where the
factor $2\le e^{\ell_s}$ is absorbed by lowering the exponent by one (valid since $\ell_s\ge 1$, hence $2\le e\le e^{\ell_s}$).

\emph{Sparse case $\mu_s<\ell_s$.} Here the relevant overload scale is the additive deviation $\Psifn(\mu_s,\ell_s)$, so we
use a hard threshold. Let $Y_i=\1[r_i>K\Psifn(\mu_s,\ell_s)]$, again nondecreasing in $r_i$. Its expectation is precisely
the probability that the binomial load $r_i\sim\Bin(s,d/n)$ exceeds $t=K\Psifn(\mu_s,\ell_s)$. Bennett's inequality for a sum
of independent indicators of mean $\mu_s$ states $\Pr(r_i\ge\mu_s+x)\le\exp(-\mu_s\,h(x/\mu_s))$ with
$h(u)=(1+u)\log(1+u)-u$, and writing $x=t-\mu_s$ this is the upper tail at the deviation $t$ above the mean. Since
$t=K\Psifn\ge K\sqrt{\mu_s\ell_s}\ge K\mu_s$ (using $\ell_s>\mu_s$ in this case), we have $x=t-\mu_s\ge t/2$ once $K\ge 2$,
so the exponent is at least $\tfrac12\,\mu_s\,h(t/(2\mu_s))$ up to the relabelling $x\ge t/2$; we record the two sub-regimes
governed by $\ell_s/\mu_s$ explicitly. There are two sub-regimes governed by $\ell_s/\mu_s$. When $\ell_s/\mu_s=\Oh(1)$ the
deviation $t$ is comparable to $\sqrt{\mu_s\ell_s}$ (the first two summands of $\Psifn$ being of the same order there), so
$t\lesssim\mu_s$ up to constants, the Gaussian branch $h(u)=\Th(u^2)$ of Bennett's exponent applies, and it contributes
\[
  \mu_s\,h(t/\mu_s)\ \gtrsim\ \mu_s\cdot\frac{t^2}{\mu_s^2}\ =\ \frac{t^2}{\mu_s}\ \ge\ \frac{K^2\mu_s\ell_s}{\mu_s}\ =\ K^2\ell_s.
\]
When $\ell_s/\mu_s\to\infty$ the threshold is dominated by the Poissonian summand $K\ell_s/\log(e+\ell_s/\mu_s)$ of $\Psifn$,
so $t\gtrsim\mu_s$, the large-deviation branch $h(u)=\Th(u\log(1+u))$ applies, and
\[
  \mu_s\,h(t/\mu_s)\ \gtrsim\ \mu_s\cdot\frac{t}{\mu_s}\log\!\Big(1+\frac t{\mu_s}\Big)\ =\ t\,\log\!\Big(1+\frac t{\mu_s}\Big)
  \ \gtrsim\ \frac{K\ell_s}{\log(e+\ell_s/\mu_s)}\cdot\log\!\Big(\frac t{\mu_s}\Big)\ \gtrsim\ K\ell_s,
\]
because $t/\mu_s\gtrsim(\ell_s/\mu_s)/\log(e+\ell_s/\mu_s)$ makes $\log(t/\mu_s)\asymp\log(\ell_s/\mu_s)\asymp\log(e+\ell_s/\mu_s)$,
so the amplification $\log(t/\mu_s)$ is exactly cancelled by the denominator $\log(e+\ell_s/\mu_s)$ built into $\Psifn$; the
additive ``$1$'' inside $\Psifn$ keeps $t\ge K$ and the bound meaningful as $\mu_s\to0$. In both sub-regimes the exponent is
at least $B\ell_s$ once $K$ is large in terms of $B$ (take $K^2\ge B$ in the first and $cK\ge B$ in the second), so
$\E Y_i=\Pr(r_i>K\Psifn)\le e^{-B\ell_s}$.

\paragraph{From expectation to high probability, and the union.}
We now turn the per-row expectation bound into control of the number of overloaded rows, and then pay for all subsets. In
both cases $Y_i$ is a nondecreasing function of $r_i$, hence so is $e^{\theta Y_i}$ for $\theta\ge 0$, and since the rows
$\{r_i(S)\}_i$ are NA the upper-tail factorization of \Cref{sec:NA} gives, for every $\theta\ge0$,
\[
  \E\Big[\exp\!\Big(\theta\sum_i Y_i\Big)\Big]\ =\ \E\Big[\prod_i e^{\theta Y_i}\Big]\ \le\ \prod_i\E\big[e^{\theta Y_i}\big],
\]
exactly as under independence. The count $\sum_i Y_i$ has mean $M=\sum_i\E Y_i\le n e^{-B\ell_s}$, and we ask that it not
exceed $a=s/64$. Applying Markov's inequality to the factorized moment generating function and optimizing $\theta$ in the
usual way is the standard Chernoff bound for a sum of $[0,1]$ variables, which with this factorization gives
\[
  \Pr\Big(\sum_i Y_i>a\Big)\ \le\ \Big(\tfrac{eM}{a}\Big)^{a}\ \le\ \big(64e\,(n/s)\,e^{-B\ell_s}\big)^{s/64}\ \le\ e^{-cB\,s\ell_s},
\]
where the middle step substitutes $M\le n e^{-B\ell_s}$ and $a=s/64$, and the last step uses $\log(n/s)\le\ell_s+\Oh(1)$
(which holds because $T=\Th(n)$, so $\log(n/s)$ and $\ell_s=\log(eT/s)$ differ by $\Oh(1)$); the factor $64e\,(n/s)$ is thus
$e^{\Oh(\ell_s)}$ and is swallowed by the dominant $e^{-Bs\ell_s/64}$, leaving an exponent $cB\,s\ell_s$ for an absolute
constant $c$. This says that for a \emph{fixed} $S$ the probability that more than $s/64$ rows are overloaded is at most
$e^{-cB\,s\ell_s}$. It remains to union over $S$. The number of $s$-subsets is $\binom{T}{s}\le(eT/s)^s=e^{s\ell_s}$, so
summing the failure probability over all subsets of each size $s=1,\dots,T$,
\[
  \Pr(\text{some }S\text{ fails})\ \le\ \sum_{s=1}^{T} e^{s\ell_s}\,e^{-cB\,s\ell_s}\ =\ \sum_{s=1}^{T} e^{-(cB-1)s\ell_s}.
\]
Here the enumeration cost $e^{s\ell_s}$ and the per-subset bound $e^{-cB\,s\ell_s}$ carry the \emph{same} weight $s\ell_s$,
which is what makes a single constant $B$ suffice across all scales. The function $s\ell_s=s\log(eT/s)$ is increasing on
$[1,T]$, so $s\ell_s\ge\ell_1=\log(eT)$ for every $s$; hence each term is at most $(eT)^{-(cB-1)}$, and the sum of at most
$T$ such terms is at most $T(eT)^{-(cB-1)}$. This is $o(1)$ once $cB>2$, which fixes the choice of $B$, and the constant $K$
is the $K(B)$ produced by the tail estimates above. \qed

\subsection{The splitting lemma (\Cref{lem:split})}\label{app:split-full}

The goal of the splitting lemma is to halve a column set $S$ into a part $Y$ of size $a$ so that every row load is split
almost perfectly in proportion, with only an additive $\Psifn$ error. We obtain $Y$ by partial coloring: we run the
Lovett--Meka walk, which finds a sign vector that lands on the boundary of a body cut out by a handful of linear
constraints, and we tune those constraints, one per row, so that the rows whose loads we must control are kept on
near-equality faces while the rest are free to move. The size constraint is enforced by one extra equality. The walk is
then summed over $\Oh(\log m)$ halving stages, and the per-stage errors form a geometric series that collapses to a single
$\Psifn$ term.

Run \Cref{lem:LM} from the starting point $x^{(0)}_t=2a/s-1$ ($t\in S$), so that the target average sign encodes the
intended split size: the entries of $x^{(0)}$ average to $\tfrac1s\sum_t(2a/s-1)=2a/s-1$, which is the mean sign of a set of
size $a$. A sign vector $\sigma\in\{-1,1\}^S$ with $\sum_t\sigma_t=2a-s$ corresponds to the set $Y=\{t:\sigma_t=1\}$ of size
$|Y|=a$, since if $|Y|=a$ then $\sum_t\sigma_t=a\cdot(+1)+(s-a)\cdot(-1)=2a-s$, and conversely. To pin the cardinality we
always include the all-ones constraint vector $v_0=\mathbf 1$ with threshold $c_0=0$. By \Cref{lem:LM} the threshold $0$
enforces the exact linear equality $\langle\mathbf 1,\sigma-x^{(0)}\rangle=0$, that is $\sum_t\sigma_t=\sum_t x^{(0)}_t=2a-s$,
so the average sign is preserved exactly from the start point through the end of the walk. This is what guarantees $|Y|=a$ on
exit, not merely in expectation. Let $A$
denote the current alive set of columns, $m=|A|$, and write $\mu_A=dm/n$, $\ell_A=\log(eT/m)$ for the running mean and
logarithmic budget. The Lovett--Meka guarantee is that the walk reaches a face on which at least a constant fraction of the
coordinates are frozen to $\pm1$, provided the constraints leave enough slack; the constraints we impose are exactly the
overload tests of the structural lemma, with locally-tuned thresholds, and we verify the slack in each phase.

\emph{Dense phase ($\mu_A\ge\ell_A$).} In this phase we constrain every nonempty row by a soft Gaussian-width budget scaled
to its own load. For row $i$ feed \Cref{lem:LM} the test vector $v_i\in\R^A$ with $(v_i)_t=A_{it}$ for $t\in A$, the
indicator of which alive columns cover $i$, so that $\langle v_i,\sigma\rangle=\sum_{t\in A}A_{it}\sigma_t$ is exactly the
signed load of row $i$ under the partial coloring and $\|v_i\|_2=\sqrt{\sum_{t\in A}A_{it}}=\sqrt{r_i(A)}$. For each row with
$r_i(A)>0$ set the threshold coefficient $c_i=K\sqrt{\mu_A\ell_A}/\sqrt{r_i(A)}$, so that heavily loaded rows are held more
tightly. The Lovett--Meka feasibility condition is that $\sum_i e^{-c_i^2/16}$ be a small fraction of $m$. Substituting
$c_i^2=K^2\mu_A\ell_A/r_i(A)$,
\[
  \sum_{i:\,r_i(A)>0} e^{-c_i^2/16}\ =\ \sum_{i:\,r_i(A)>0}\exp\!\Big(-\frac{K^2\mu_A\ell_A}{16\,r_i(A)}\Big)\ \le\ \frac{m}{64}
\]
is precisely the dense structural estimate \eqref{eq:struct-dense} read at the alive set $A$ (with $s=m$, $\mu_s=\mu_A$,
$\ell_s=\ell_A$). This leaves room for the one extra all-ones equality constraint, which we encode as the vector
$\mathbf 1\in\R^A$ with threshold $0$ (contributing $e^{-0}=1$ to the budget): the total is then
\[
  \sum_i e^{-c_i^2/16}+e^{-0}\ \le\ \frac{m}{64}+1\ \le\ \frac{m}{16}
\]
as soon as $m\ge 22$ (where $m/16-m/64=3m/64\ge1$), so the total constrained budget stays below the threshold the walk
requires by \Cref{lem:LM}. Each iteration moves $\sigma$ along a unit Gaussian step inside the body, so the guarantee
$|\langle v_i,\sigma-x_0\rangle|\le c_i\|v_i\|_2$ bounds the change in the load of row $i$ by
\[
  c_i\,\|v_i\|_2\ =\ \frac{K\sqrt{\mu_A\ell_A}}{\sqrt{r_i(A)}}\cdot\sqrt{r_i(A)}\ =\ K\sqrt{\mu_A\ell_A},
\]
where the $\sqrt{r_i(A)}$ in $\|v_i\|_2$ cancels the $\sqrt{r_i(A)}$ in the denominator of $c_i$. The per-row, per-stage
change is therefore the uniform bound $K\sqrt{\mu_A\ell_A}$, independent of $i$.

\emph{Sparse phase ($\mu_A<\ell_A$).} Here loads are small and we use a hard threshold. Set $U_A=K\Psifn(\mu_A,\ell_A)$. By
the sparse structural estimate \eqref{eq:struct-sparse} read at $A$, the overloaded set $\mathcal O=\{i:r_i(A)>U_A\}$ has
$|\mathcal O|\le m/64$. We constrain exactly those overloaded rows: for each $i\in\mathcal O$ feed the same load vector
$v_i=(A_{it})_{t\in A}$ with threshold $c_i=0$, which by \Cref{lem:LM} freezes $\langle v_i,\sigma-x^{(0)}\rangle=0$ so the
walk does not move row $i$'s load at all. Including the all-ones equality vector, the budget is
\[
  \sum_{i\in\mathcal O} e^{-0}+e^{-0}\ =\ |\mathcal O|+1\ \le\ \frac{m}{64}+1\ \le\ \frac{m}{16}
\]
for $m\ge22$, so \Cref{lem:LM} again applies. We \emph{release} all the remaining rows, imposing no constraint on them. A released row had at most $U_A$ alive incidences to begin with, so no matter how its
incidences are colored across this and all later stages, its load can change by at most $U_A$ in each of the two halves,
hence by at most $2U_A$ in total. Releasing such rows costs nothing because their entire future deviation is already bounded
by $2U_A=\Oh(\Psifn(\mu_A,\ell_A))$.

\emph{Stage-summing.} It remains to add the per-stage changes along the whole recursion and check that the total is still
$\Oh(\Psifn(\mu_0,\ell_0))$, with no logarithmic blow-up from the number of stages. The alive set halves at each stage, so
after $u$ natural-log steps of halving the mean has shrunk geometrically; writing the shrink factor as $e^{-u}$ we have
$\mu(u)=\mu_0 e^{-u}$ and $\ell(u)=\ell_0+u$, since halving the column count adds a bounded amount to the logarithm at each
stage. For the dense stages, the per-stage change is $K\sqrt{\mu_A\ell_A}=K\sqrt{\mu(u)\ell(u)}=K e^{-u/2}\sqrt{\ell_0+u}$.
The function $e^{-u/2}\sqrt{\ell_0+u}$ is nonincreasing in $u$ for $\ell_0\ge 1$ (the exponential decay dominates the slow
growth of the square root), so summing it over the discrete halving stages is bounded by its value at the start times a
geometric factor, giving $\sum_{\text{dense}}\sqrt{\mu_A\ell_A}\le C\sqrt{\mu_0\ell_0}$. The geometric decay is what
prevents a factor of the number of stages from appearing. For the sparse thresholds, what we need is
$\sup_{\text{sparse}}\Psifn(\mu_A,\ell_A)\le C\Psifn(\mu_0,\ell_0)$, since a released row's contribution is $\Oh$ of a single
such threshold. If $\mu_0\le\ell_0$ then the whole recursion is sparse and $\Psifn(\mu_A,\ell_A)$ is monotone in the stage,
so the supremum is attained at the start and equals $\Psifn(\mu_0,\ell_0)$. If $\mu_0>\ell_0$, the recursion begins dense
and only enters the sparse phase once $\mu(u)$ has dropped below $\ell(u)$, which happens at some stage
$u_*\le\log(\mu_0/\ell_0)$; at that point $\ell_*=\ell_0+u_*\le\ell_0+\log(\mu_0/\ell_0)\le C\sqrt{\mu_0\ell_0}$, so the
sparse phase effectively restarts from a budget already controlled by $\sqrt{\mu_0\ell_0}\le C\Psifn(\mu_0,\ell_0)$, and we
rebase the sparse bound there. The recursion stops once $\Oh(1)$ variables remain alive, at which point the residual
coloring is found by brute force at cost $\Oh(1)$ per row; this constant is absorbed by the additive ``$1$'' inside
$\Psifn$. Finally, the cardinality is exact up to an $\Oh(1)$ cleanup swap to repair any unit imbalance, and such a swap
changes each $r_i(Y)$ by at most $1\le\Psifn$. Dividing the resulting signed discrepancy by $2$ converts the symmetric
$\pm1$ split into the two-sided load bound claimed. \qed

\subsection{The recursive color-tree (\Cref{lem:offline-upper})}\label{app:tree-full}

We build the $k$-coloring by repeatedly halving the column set with \Cref{lem:split}, forming a balanced binary tree whose
leaves are the $k$ color classes. The key point is bookkeeping: each split introduces a fresh additive error into the row
loads, but that error is then carried down to the leaf scaled by the fraction of columns that survive, and these fractions
multiply to a small number. We track one row $i$ and one root-to-leaf path, write the leaf load as the ideal proportional
load plus the accumulated error, and show the accumulated error sums to a single $\Psifn$ with no extra logarithmic factor
from the depth of the tree.

For a node $v$ with $q_v$ leaves below it and column set of size $M_v$, applying \Cref{lem:split} produces children $w$
whose loads satisfy $r_i(w)=(M_w/M_v)r_i(v)+e_i(w)$, where the first term is the perfectly proportional split and the
additive error obeys $|e_i(w)|\le C\Psifn(dM_v/n,\log(eT/M_v))$ by the splitting lemma. We index nodes by the number $q$ of
leaves they subtend. For a $q$-leaf node the relevant parameters are $dM_v/n=\Th(\lam q)$, since such a node holds a
$\Th(q/k)$ fraction of all columns and $\lam=\Th(dk/n)/k$, and $\log(eT/M_v)=\Th(1+\log(k/q))=:\ell_q$. So the error
injected at a $q$-node is $\Oh(\Psifn(\Th(\lam q),\ell_q))$.

Now follow this error to the leaf. Passing the recursion $r_i(w)=(M_w/M_v)r_i(v)+e_i(w)$ down the path expresses the leaf
load $x_{i,c}$ as the ideal load $\tfrac{m_c}{T}\deg(i)$ plus a sum of the injected errors, each multiplied by the product
of the child-fractions $M_{(\cdot)}/M_{(\cdot)}$ on the remaining path from that node to the leaf. Along a balanced path
those fractions are each $\Th(1/2)$ and they telescope: the product of the fractions from a $q$-node down to the leaf is the
ratio of the leaf's column count to the $q$-node's column count, which is $\Th(1/q)$. Hence the error born at a $q$-node
reaches the leaf attenuated by $\Th(1/q)$. Summing the attenuated errors along the root-to-leaf path,
\[
  \Big|x_{i,c}-\tfrac{m_c}{T}\deg(i)\Big|\ \le\ C\sum_{q\text{ on path}}\frac{\Psifn(\Th(\lam q),\ell_q)}{q}
  \ =\ C\Big(\underbrace{\sum_q\tfrac1q}_{\Oh(1)}+\sqrt\lam\sum_q\sqrt{\tfrac{\ell_q}{q}}+\sum_q\frac{\ell_q}{q\,\log(e+\ell_q/(\lam q))}\Big),
\]
where the three groups come from expanding $\Psifn(\Th(\lam q),\ell_q)=\Th(1)+\Th(\sqrt{\lam q\,\ell_q})+\Th(\lam
q\,\ell_q/(\lam q\log(e+\ell_q/(\lam q))))$ and dividing each piece by $q$: the constant ``$1$'' inside $\Psifn$ gives
$\sum_q 1/q$, the Gaussian part $\sqrt{\lam q\,\ell_q}/q=\sqrt\lam\sqrt{\ell_q/q}$ gives the middle sum, and the Bennett part
$(\lam q)\ell_q/(\lam q\cdot\log(e+\ell_q/(\lam q)))$ divided by $q$ gives the last sum after cancelling $\lam q$. The
values of $q$ along the path are the geometric sequence $k,k/2,\dots,2$, so each of the three sums is a geometric-type sum
that we now evaluate.

The first sum is harmless. Writing $q=k/2^j$ for $j=0,1,\dots,\log_2 k-1$, the values of $1/q$ form the geometric
progression $1/k,2/k,4/k,\dots,1/2$, so
\[
  \sum_{q} \frac1q\ =\ \sum_{j=0}^{\log_2 k-1}\frac{2^j}{k}\ =\ \frac{2^{\log_2 k}-1}{k}\ =\ \frac{k-1}{k}\ \le\ 1\ =\ \Oh(1).
\]

\emph{Square-root term.} We must bound $\sum_q\sqrt{\ell_q/q}$ with $\ell_q=1+\log(k/q)$. Substitute $q=k/2^j$, so that
$\ell_q=1+\log 2^j=1+j\log 2=\Th(1+j)$ and $1/q=2^j/k$, giving the explicit series
\[
  \sum_{q}\sqrt{\frac{\ell_q}{q}}\ =\ \sum_{j=0}^{\log_2 k-1}\sqrt{\frac{(1+j\log 2)\,2^j}{k}}\ =\ \frac1{\sqrt k}\sum_{j=0}^{\log_2 k-1}\sqrt{1+j\log 2}\;\,2^{j/2}.
\]
The summand grows by the geometric factor $2^{1/2}$ times the slowly varying $\sqrt{(1+(j+1)\log2)/(1+j\log2)}\to1$, so the
series is geometric and dominated by its last term $j=\log_2 k-1$, where $1+j\log2=\Th(\log k)=\Th(\Lscale)$ and
$2^{j/2}=\Th(\sqrt k)$. Hence
\[
  \sum_{q}\sqrt{\frac{\ell_q}{q}}\ \le\ \frac{C}{\sqrt k}\cdot\sqrt{\Lscale}\cdot\sqrt k\ =\ \Oh\big(\sqrt{\Lscale}\big),
\]
the geometric tail summing to a constant multiple of the largest term. Multiplying by the prefactor $\sqrt\lam$, the whole
square-root term is $\Oh(\sqrt{\lam\Lscale})$.

\emph{Bennett term.} We must bound $\sum_q \ell_q/\big(q\,\log(e+\ell_q/(\lam q))\big)$. We first record the geometric sum
that appears repeatedly: by the same substitution $q=k/2^j$ used above, $\sum_{q\le Q}\ell_q/q$ is dominated by its
largest-$j$ (smallest-$q$) term, where $\ell_q=\Th(\log(k/q))$ peaks and $1/q$ peaks, so for any cutoff $Q$,
\[
  \sum_{q\le Q}\frac{\ell_q}{q}\ \le\ C\,\frac{\ell_Q}{Q}\cdot\frac{1}{1-1/2}\ =\ \Oh\!\Big(\frac{\ell_Q}{Q}\Big),
  \qquad\text{and in particular}\qquad \sum_{q}\frac{\ell_q}{q}\ =\ \Oh(\Lscale)
\]
taking $Q=2$ at the small-$q$ end where $\ell_q=\Th(\log k)=\Th(\Lscale)$ and $1/q=\Th(1)$. Now split into cases.

If $\lam\ge\Lscale$, then for every $q\ge1$ we have $\ell_q/(\lam q)\le\ell_q/\lam\le\Lscale/\lam\le 1$, so the denominator
$\log(e+\ell_q/(\lam q))\ge\log e=1$, and dropping it,
\[
  \sum_q \frac{\ell_q}{q\,\log(e+\ell_q/(\lam q))}\ \le\ \sum_q\frac{\ell_q}{q}\ =\ \Oh(\Lscale)\ =\ \Oh(\sqrt{\lam\Lscale}),
\]
the last equality holding because $\Lscale\le\sqrt{\lam\Lscale}$ when $\lam\ge\Lscale$.

If $\lam<\Lscale$, put $A=\Lscale/\lam>1$ and split the path at $q=\sqrt A$. For the small-$q$ part $q\le\sqrt A$ we have
$\ell_q\asymp\Lscale$ and $\ell_q/(\lam q)\ge\Lscale/(\lam\sqrt A)=A/\sqrt A=\sqrt A$, so
$\log(e+\ell_q/(\lam q))\ge\log(e+\sqrt A)\ge c\log(e+A)$, and pulling this denominator out of the geometric sum,
\[
  \sum_{q\le\sqrt A}\frac{\ell_q}{q\,\log(e+\ell_q/(\lam q))}\ \le\ \frac{1}{c\log(e+A)}\sum_{q\le\sqrt A}\frac{\ell_q}{q}
  \ =\ \Oh\!\Big(\frac{\Lscale}{\log(e+A)}\Big).
\]
For the large-$q$ part $q>\sqrt A$ we simply drop the denominator (it is $\ge1$) and use the geometric sum from the other
end, now dominated by its smallest surviving $q=\Th(\sqrt A)$ where $\ell_q=\Th(\Lscale)$:
\[
  \sum_{q>\sqrt A}\frac{\ell_q}{q}\ \le\ C\,\frac{\Lscale}{\sqrt A}\ =\ C\,\frac{\Lscale}{\sqrt{\Lscale/\lam}}\ =\ C\sqrt{\lam\Lscale}.
\]
Adding the two parts, the Bennett term is $\Oh(\sqrt{\lam\Lscale}+\Lscale/\log(e+\Lscale/\lam))$.

Adding the three pieces, $\Oh(1)+\Oh(\sqrt{\lam\Lscale})+\Oh(\sqrt{\lam\Lscale}+\Lscale/\log(e+\Lscale/\lam))$, recombines
exactly into $\Oh(\Psifn(\lam,\Lscale))$ with $\Lscale=\log(ek)$; no extra logarithmic factor survives because the
attenuation $1/q$ turned each path sum into a geometric series dominated by one endpoint. Finally we pass from per-color
loads to the range. The ideal loads of two colors differ by
$|(m_c/T)\deg(i)-(m_{c'}/T)\deg(i)|\le\deg(i)/T\le 1$, because the class sizes $m_c$ differ by at most one and
$\deg(i)\le T$; therefore $\rangeop_i\le 2\max_c|x_{i,c}-(m_c/T)\deg(i)|+1=\Oh(\Psifn(\lam,\log k))$ at every coordinate
$i$. The whole construction is deterministic once the incidence matrix $A$ is fixed, because the structural event of
\Cref{lem:structural} simultaneously covers every alive set the recursion can select, so no further randomness is spent
beyond drawing $A$. \qed
