% appC-online.tex -- full deferred proofs of Section 4.
\section{Full proof of the online bounds}\label{app:online}

\subsection{The lower bound (\Cref{lem:online-lower})}\label{app:online-lower-full}

We transfer the two-color online lower bound to $k$ colors by projecting the colors onto two through a fixed zero-sum signing,
and prove the resulting two-action lower bound from scratch, so the online lower bound is self-contained.

\paragraph{Projection.}
Fix a signing $\xi:[k]\to\R$ with $\sum_c\xi(c)=0$ and apply it coordinatewise to an online $k$-coloring: the value
$\xi(\chi(t))$ is a function of the online color $\chi(t)$, hence itself an online decision. The signed sum at coordinate $i$
is $D_i(t)=\sum_{s\le t:\,i\in S_s}\xi(\chi(s))=\sum_c\xi(c)x_{i,c}(t)$. For even $k$ split $[k]$ into two halves with
$\xi\equiv+1$ and $\xi\equiv-1$; for odd $k=2a+1$ put $\xi\equiv+1$ on a part of size $a$ and $\xi\equiv-\tfrac{a}{a+1}$ on a
part of size $a+1$. In both cases the total positive weight equals the total negative weight, call it $w$ (with $w\le k/2$),
so pairing positive against negative terms gives $|D_i(t)|\le w\,\rangeop_i(t)$; and the induced online process takes steps in
$\{+1,-\rho\}$ with $\rho=1$ for even $k$ and $\rho=\tfrac{a}{a+1}\in[\tfrac12,1]$ for odd $k$. The lower bound follows from
the next lemma, applied to the projected process.

\begin{lemma}[Asymmetric two-action online lower bound]\label{lem:asym-online}
Fix $\rho\in[\tfrac12,1]$. Let $S_1,\dots,S_T$ be i.i.d.\ uniform $d$-subsets of $[n]$ with $\tau_-n\le T\le\tau_+n$ and
$2\le d\le\Hscale^2/\log\Hscale$. For any online rule choosing $\eta_t\in\{+1,-\rho\}$ from $S_t$ and the past, set
$D_i(t)=\sum_{s\le t:\,i\in S_s}\eta_s$. Then $\Pr\big(\max_{t\le T,\,i\le n}|D_i(t)|\ge c\,\Hscale\big)=1-o(1)$ for an
absolute $c>0$, uniformly in $\rho\in[\tfrac12,1]$.
\end{lemma}

\begin{proof}
Conditioning on the rule's auxiliary randomness, we may assume it is deterministic and adapted to the revealed supports.
Write $L=\log n$.

\emph{Fresh-block bridge estimate.} Suppose at a history-measurable time $K$ there are disjoint $I^-,I^+\subseteq[n]$ with
$|I^-|=|I^+|=s$, and the next $\Delta$ supports are fresh, with $n^{0.99}\le s\le n/L$ and $n^{0.99}\le\Delta\le n/L^2$.
We claim that, conditionally on the history through $K$, with probability $1-o(1/L)$ both: (1) at least $s/2$ vertices of each
of $I^-,I^+$ are untouched in the block; and (2) there are at least $c_0\Delta(s/n)^2$ block columns $t$ with
$|S_t\cap I^-|=|S_t\cap I^+|=1$ whose two chosen vertices appear in no other block column, for an absolute $c_0>0$.

For (1): for $i\in I^\pm$ let $Z_i=\1[i\text{ untouched}]$, so $\E Z_i=(1-d/n)^\Delta=1-o(1)$ (as $d\Delta/n\le d/L^2=o(1)$),
and for $i\ne i'$, $\E[Z_iZ_{i'}]=\mu^2(1+\Oh(d\Delta/n^2))$; hence $\Var(\sum_{I^\pm}Z_i)\le s+\Oh(s^2 d\Delta/n^2)=o(s^2)$
and Chebyshev gives $\Pr(\sum Z_i<s/2)=\Oh(1/s+d\Delta/n^2)=o(1/L)$. For (2): for a column $t$ let $Y_t$ indicate it is an
isolated bridge; counting gives $\E Y_t=(1+o(1))d(d-1)(s/n)^2\ge(s/n)^2=:\nu$ (using $ds/n,d\Delta/n=o(1)$), so $\E W=\Delta\nu\ge\Delta(s/n)^2$
for $W=\sum_t Y_t$. For the second moment, classify pairs of distinct block columns $t\ne t'$ by how their chosen endpoints
and supports interact. (a) If the four endpoints $i_t^-,i_t^+,i_{t'}^-,i_{t'}^+$ are distinct and the two supports meet
neither the others' endpoints nor $I^-\cup I^+$ again, the two events are almost independent: $\E[Y_tY_{t'}]=\nu^2(1+\Oh(d^2\Delta/n^2))$,
the correction being the chance that one of the $\le d$ further slots of $S_t$ or $S_{t'}$ lands in the $\Oh(s)$ relevant
coordinates. (b) If $t,t'$ share an endpoint (probability $\Oh(1/s)$ given both are bridges, since the shared vertex is one of
$\asymp s$ equally likely choices), the contribution is $\Oh(\nu^2/s)$ after summing. There is no contribution from a column
hitting $I^\pm$ twice, as that is excluded by the isolation requirement. Summing over the $\Delta(\Delta-1)$ ordered pairs and
adding the diagonal $\E W$,
\[
  \Var W\ =\ \E W+\sum_{t\ne t'}\big(\E[Y_tY_{t'}]-\nu^2\big)\ \le\ \E W+\Delta^2\nu^2\,\Oh\!\Big(\tfrac1s+\tfrac{d^2\Delta}{n^2}\Big),
\]
so $\Var W/(\E W)^2\le 1/\E W+\Oh(1/s+d^2\Delta/n^2)=\Oh\big(1/(\Delta(s/n)^2)+1/s+d^2\Delta/n^2\big)=o(1/L)$, and Chebyshev
gives $W\ge\tfrac12\E W$ with probability $1-o(1/L)$. Neither estimate inspects the chosen steps; both use only fresh-support
independence.

\emph{Real-valued spreads.} Say the process has an \emph{$(\ell,r)$-spread of size $s$} at time $K$ if there are disjoint
$I^-,I^+$ of size $s$ with $D_i(K)=\ell\le0$ for $i\in I^-$ and $D_i(K)=r\ge0$ for $i\in I^+$ ($\ell,r\in\R$). Set $C=8$,
\[
  q_*=\Big\lfloor\tfrac{\Hscale}{2\log C}\Big\rfloor,\quad
  s_q=\big\lfloor n/L^{C^q}\big\rfloor,\quad
  \Delta_q=\big\lfloor n/L^{2q}\big\rfloor,\quad K_q=\sum_{u=1}^{q}\Delta_u\ \ (K_0=0).
\]
Then $K_{q_*}\le 2n/L^2=o(n)$, so $K_{q_*}\le T$, and for $q\le q_*$ one has $C^{q_*}\le\sqrt L$, whence $s_q,\Delta_q\ge n^{0.99}$.
We show inductively that with probability $1-o(1)$, for every $0\le q\le q_*$ there is an $(\ell_q,r_q)$-spread of size
$\ge s_q$ at time $K_q$ with
\begin{equation}\label{eq:spread-grow}
  r_q-\ell_q\ \ge\ q\rho .
\end{equation}
At $q=0$, take any two disjoint $s_0$-sets; all values are $0$, so \eqref{eq:spread-grow} holds with $\ell_0=r_0=0$. Assume it
at $q-1$, on sets $I^-,I^+$ of size $s=s_{q-1}$. Apply the fresh-block estimate to the next $\Delta_q$ columns; let $J$ be the
isolated bridges, so $|J|\ge c_0\Delta_q(s/n)^2$ on its good event. For $t\in J$ with endpoints $i_t^-\in I^-$, $i_t^+\in I^+$,
exactly one of $\eta_t=+1$ (then $D_{i_t^+}(K_q)=r+1$) or $\eta_t=-\rho$ (then $D_{i_t^-}(K_q)=\ell-\rho$) holds. At least
$|J|/2$ bridges share the same alternative. If $\ge|J|/2$ choose $+1$, pair their high endpoints (at $r+1$) with untouched low
vertices (at $\ell$): the new gap is $(r+1)-\ell\ge(r-\ell)+\rho$. If $\ge|J|/2$ choose $-\rho$, pair their low endpoints (at
$\ell-\rho$) with untouched high vertices (at $r$): the new gap is $r-(\ell-\rho)=(r-\ell)+\rho$. Either way the gap grows by
$\ge\min\{1,\rho\}=\rho$. The new spread has size $\ge\min\{s/2,|J|/2\}\ge s_q$, since $s/2\ge s_q$ and
$|J|/2\ge c\,nL^{-2C^{q-1}-2q}\ge nL^{-C^q}\ge s_q$ as $C^q-2C^{q-1}-2q=6C^{q-1}-2q\ge4$. Each stage fails with probability
$o(1/L)$, and $q_*=\Oh(\Hscale)$, so a union bound completes the induction. At $q_*$, $r_{q_*}-\ell_{q_*}\ge\rho q_*$ and
$\ell_{q_*}\le0\le r_{q_*}$, so $\max_i|D_i(K_{q_*})|\ge\rho q_*/2\ge\Hscale/(16\log 8)$. This proves the lemma.
\end{proof}

Applying \Cref{lem:asym-online} to the projected process: $w\,\max_{t,i}\rangeop_i\ge\max_{t,i}|D_i(t)|\ge c\Hscale$ \whp,
and $w\le k/2$, so $\ondisc_k\ge\Om(\loglog n/k)$, for even and odd $k$ alike. For fixed $k$ this is $\Om(\loglog n)$.

\subsection{The upper bound (\Cref{lem:online-upper})}\label{app:online-upper-full}

The plan is to compare the actual coloring against the pure seed coloring, show that the seed alone keeps every range small,
and then show that the few repairs needed on top of the seed change each range by only a controlled amount. Write
$\Hscale=\loglog n$. Let $\widetilde x_{i,c}(t)$ be the loads under the pure-seed coloring ($\chi\equiv\widetilde\chi$), the
coloring obtained by always following the seed and never repairing, and let $\widetilde R_i(t)$ be the resulting seed-only
range. The two colorings agree except where the algorithm steps in, and we measure that gap by counting interventions. A
coordinate--arrival pair $(i,t)$ is a \emph{modified incidence} if $i\in S_t$ and the algorithm's color differs from
$\widetilde\chi_t$; by construction the color differs only when the repair branch fires, and that branch requires a
coordinate $i^*\in S_t\cap E_{t-1}$ to repair. Each modified incidence at $i$ moves exactly one of the load coordinates of
$x_i$: the seed would have added $1$ to color $\widetilde\chi_t$, while the algorithm adds $1$ to its repair color instead,
so the two load vectors differ by $+1$ in one coordinate and $-1$ in another at that step. Each such step changes
$x_{i,c}(t)-\widetilde x_{i,c}(t)$ by at most $1$ in absolute value in each color $c$, so accumulating over steps,
\[
  \|x_i(t)-\widetilde x_i(t)\|_\infty\ \le\ \#\{\text{modified incidences at }i\text{ up to time }t\}.
\]
The range $R_i=\max_c x_{i,c}-\min_c x_{i,c}$ is $2$-Lipschitz in the $\ell_\infty$ norm of the load vector, because
$|\max_c x_{i,c}-\max_c\widetilde x_{i,c}|\le\|x_i-\widetilde x_i\|_\infty$ and likewise for the min, and the range is their
difference. Hence
\[
  |R_i(t)-\widetilde R_i(t)|\ \le\ 2\,\|x_i(t)-\widetilde x_i(t)\|_\infty\ \le\ 2\cdot(\#\text{ modified incidences at }i).
\]
So once we bound the seed range and the number of modified incidences, we bound the actual range.

\paragraph{Part 1: seed tail.}
The first job is to show that the seed coloring by itself almost never lets any coordinate's range reach $32\Hscale$, so the
repairs only ever have to deal with a sparse set of coordinates. The range at $i$ is a maximum over color pairs of the gap
between two loads, so we track each pair separately. Fix a coordinate $i$ and two colors $a\ne b$, and let
$$\widetilde D_{i,a,b}(t)=\sum_{s\le t}\1[i\in S_s]\big(\1[\widetilde\chi_s=a]-\1[\widetilde\chi_s=b]\big),$$
which is the running difference between the load of color $a$ and the load of color $b$ at coordinate $i$ under the seed.
Because the seed colors $\widetilde\chi_s$ are independent across $s$ and the increment at step $s$,
$\Delta_s=\1[i\in S_s](\1[\widetilde\chi_s=a]-\1[\widetilde\chi_s=b])$, has conditional mean zero (the seed assigns $a$ and
$b$ with equal probability $1/k$, so $\E[\1[\widetilde\chi_s=a]-\1[\widetilde\chi_s=b]\mid i\in S_s]=1/k-1/k=0$), this is a
martingale with increments $\Delta_s\in\{-1,0,1\}$. Its predictable quadratic variation is bounded by summing the per-step
conditional second moments. Each step contributes $\Delta_s^2=\1[i\in S_s]\1[\widetilde\chi_s\in\{a,b\}]$ (the cross term
vanishes since a color cannot equal both $a$ and $b$), and $i\in S_s$ and $\widetilde\chi_s\in\{a,b\}$ are independent, so
\[
  \langle\widetilde D\rangle_T\ =\ \sum_{s\le T}\E[\Delta_s^2]\ =\ \sum_{s\le T}\Pr(i\in S_s)\,\Pr(\widetilde\chi_s\in\{a,b\})
  \ =\ T\cdot\frac dn\cdot\frac2k\ =\ \frac{2Td}{nk}\ =:\ V,
\]
using that $i$ lies in a random $d$-sparse support with probability $d/n$ and that $\Pr(\widetilde\chi_s\in\{a,b\})=2/k$.
Freedman's inequality for a martingale with increments bounded by $1$ and predictable variation at most $V$ states
$\Pr(\max_t\widetilde D_{i,a,b}\ge u)\le\exp(-\tfrac{u^2/2}{V+u/3})$; taking the deviation
$u=32\Hscale$ gives
\[
  \Pr\big(\max_t\widetilde D_{i,a,b}\ge 32\Hscale\big)\ \le\ \exp\!\Big(-\frac{(32\Hscale)^2/2}{V+32\Hscale/3}\Big).
\]
The seed range at $i$ exceeds $32\Hscale$ only if some ordered pair of colors $(a,b)$ has $\widetilde D_{i,a,b}\ge 32\Hscale$,
so a union bound over the at most $k^2$ pairs gives
\[
  \Pr\big(\max_t\widetilde R_i\ge 32\Hscale\big)\ \le\ k^2\exp\!\Big(-\frac{512\Hscale^2}{V+\Oh(\Hscale)}\Big),
  \qquad 512\Hscale^2=\frac{(32\Hscale)^2}{2}.
\]
It remains to check that this bound clears $d^{-28}$ for the relevant $(k,d)$. The subtlety is that the Freedman denominator
is $V+\Oh(\Hscale)$, so the exponent does not grow without bound: with $T=\Th(n)$, $V=2Td/(nk)=\Th(d/k)$, the exponent is
$\Th(\Hscale^2/(V+\Hscale))$, which equals $\Th(\Hscale^2/V)=\Th(\Hscale^2 k/d)$ only while $V\gtrsim\Hscale$, and
\emph{saturates} at $\Th(\Hscale)$ once $V\lesssim\Hscale$. We split on this.

\emph{Regime $d/k\ge\Hscale$ (so $V\gtrsim\Hscale$).} The exponent is $\Th(\Hscale^2 k/d)$, growing linearly in $k$, while the
prefactor is only $k^2$; so $k^2\exp(-\Th(\Hscale^2 k/d))$ is decreasing in $k$ (log-derivative $2/k-\Th(\Hscale^2/d)<0$ since
$\Hscale^2/d\ge\log\Hscale\gg1$). Its maximum is at the smallest $k$, and at $k=2$, $d=\Hscale^2/\log\Hscale$ the exponent is
$\frac{512\Hscale^2}{V+\Oh(\Hscale)}=(512-o(1))\log\Hscale$, giving $\Hscale^{-512+o(1)}$. Since $d^{-28}=\Hscale^{-56+o(1)}$
here, this clears $d^{-28}$ with a factor-$9$ margin.

\emph{Regime $d/k<\Hscale$ (so $V\lesssim\Hscale$).} The exponent saturates at $\Th(512\Hscale^2/\Hscale)=\Th(\Hscale)$, so
the bound is $k^2 e^{-\Th(\Hscale)}$. For $k\le d$ this is $\le d^2 e^{-\Th(\Hscale)}=\Hscale^{\Oh(1)}(\log n)^{-\Th(1)}=o(d^{-28})$,
using $d^2\le\Hscale^{4+o(1)}$ and $e^{-\Th(\Hscale)}=(\log n)^{-\Th(1)}$. For a \emph{fixed} $k$ (any $d$), $k^2=\Oh(1)$ and
the same $e^{-\Th(\Hscale)}=(\log n)^{-\Th(1)}\le d^{-28}$ holds for large $n$ since $d^{-28}=\Om(1)$ when $d=\Oh(1)$.

So the seed tail is $\le d^{-28}$ for every $k\le d$ and every fixed $k$, which is the range of \Cref{lem:online-upper}. Thus
\begin{equation}\label{eq:seed-tail}
  \Pr(\max_t\widetilde R_i\ge 32\Hscale)\ \le\ d^{-28}.
\end{equation}

\paragraph{Part 2: a self-contained range-repair transfer lemma.}
With the seed under control, the remaining work is to show that the exceptional coordinates stay sparse and that their ranges
cannot pile up much beyond $64\Hscale$. We isolate this as a transfer lemma for any seed-and-repair rule meeting an explicit
interface, and prove it from scratch (so the upper bound does not rest on the two-color analysis as a black box). Let
$U_t\in[k]$ be the independent uniform seed colors and $C_t\in[k]$ the actual colors; $x_{i,c}(t)=\#\{s\le t:i\in S_s,C_s=c\}$,
$\rangeop_i(t)=\max_c x_{i,c}(t)-\min_c x_{i,c}(t)$, and $\widetilde x,\widetilde R$ the analogous seed quantities. Let
$E_t=\{i:\max_{s\le t}\rangeop_i(s)\ge 64\Hscale\}$.

\begin{lemma}[Range-repair transfer]\label{lem:transfer}
Assume $T\le\tau_+n$, $2\le d\le\Hscale^2/\log\Hscale$, $k\ge2$, the seed tail \eqref{eq:seed-tail}, and that $S_t$ is a
uniform $d$-subset independent of the past given $\mathcal F_{t-1}$. (The proof uses no upper bound on $k$.) Assume every non-seed assignment names a repaired
coordinate $i_t^*\in S_t\cap E_{t-1}$, and:
\begin{itemize}
  \item[\textnormal{(R1)}] $\rangeop_{i_t^*}(t)\le\rangeop_{i_t^*}(t-1)$;
  \item[\textnormal{(R2)}] $C_t\ne U_t\ \Rightarrow\ S_t\cap E_{t-1}\ne\varnothing$;
  \item[\textnormal{(R3)}] if $i\in S_t\cap E_{t-1}$ and $\rangeop_i(t)>\rangeop_i(t-1)$, there is a distinct $j\in S_t$ with
  $\rangeop_j(t-1)\ge\rangeop_i(t-1)$.
\end{itemize}
Then $\Pr\big(\max_{t\le T,\,i\le n}\rangeop_i(t)<80\Hscale\big)=1-o(1)$.
\end{lemma}

\begin{proof}
\emph{Range calculus.} For $y\in\Z_{\ge0}^k$, $\operatorname{rng}(y)=\max_c y_c-\min_c y_c$ satisfies
$|\operatorname{rng}(y+e_c)-\operatorname{rng}(y)|\le 1$ (an arrival raises any range by $\le1$) and
$|\operatorname{rng}(y)-\operatorname{rng}(z)|\le 2\|y-z\|_\infty$. Let $D_i(t)=\#\{s\le t:i\in S_s,\,C_s\ne U_s\}$; each
deviation shifts $x_i-\widetilde x_i$ by $e_{C_s}-e_{U_s}$, so $\|x_i(t)-\widetilde x_i(t)\|_\infty\le D_i(t)$ and
\begin{equation}\label{eq:range-dev}
  |\rangeop_i(t)-\widetilde R_i(t)|\le 2D_i(t).
\end{equation}
Finally, for any past-measurable $Q\subseteq[n]$ and fixed $i$,
\begin{equation}\label{eq:incidence}
  \Pr\big(i\in S_t,\ (S_t\setminus\{i\})\cap Q\ne\varnothing\,\big|\,\mathcal F_{t-1}\big)\ \le\ \frac dn\cdot\frac{(d-1)|Q|}{n-1}\ \le\ \frac{2d^2|Q|}{n^2}.
\end{equation}

\emph{Exceptional-set sparsity.} Let $M=nd^{-5}$ and $B_0=\{i:\max_{t\le T}\widetilde R_i(t)\ge 32\Hscale\}$. By
\eqref{eq:seed-tail}, $\E|B_0|\le nd^{-28}$. Since $|B_0|$ depends on the independent pairs $(S_t,U_t)$ and changing one pair
alters membership for at most $2d$ coordinates, McDiarmid's inequality \cite{McDiarmid1989} gives
\begin{equation}\label{eq:B0}
  \Pr(|B_0|>M/2)\le\exp(-c\,nd^{-12}).
\end{equation}
Let $\tau=\min\{t:|E_t|>M\}$ ($=T+1$ if never). For coordinate $i$ let $N_i$ count times $t\le T$ with $|E_{t-1}|\le M$,
$i\in S_t$, and $(S_t\setminus\{i\})\cap E_{t-1}\ne\varnothing$. Conditioning repeatedly and using \eqref{eq:incidence} with
$Q=E_{t-1}$ ($|Q|\le M$), for any $m$ prescribed times the chance all satisfy this is $\le(2d^2M/n^2)^m$, so
\begin{equation}\label{eq:Ni}
  \Pr(N_i\ge m)\le\binom Tm\Big(\frac{2d^2M}{n^2}\Big)^m\le\Big(\frac{C}{m d^3}\Big)^m .
\end{equation}
Take $m=\lfloor 15\Hscale\rfloor$ and $B_1=\{i:N_i\ge m\}$; by \eqref{eq:Ni} and Markov,
$\Pr(|B_1|>M/2)\le 2d^5(C/(md^3))^m=(\log n)^{-\omega(1)}$. Now if $\tau\le T$ and $i\in E_\tau\setminus B_0$, let $t_i\le\tau$
be its first exceptional time: $\widetilde R_i(t_i)<32\Hscale$ while $\rangeop_i(t_i)\ge 64\Hscale$, so $D_i(t_i)\ge 16\Hscale\ge m$
by \eqref{eq:range-dev}. Each of those $\ge m$ deviation-times has $i\in S_s$, $C_s\ne U_s$, hence (by (R2)) a co-present
member of $E_{s-1}$ distinct from $i$ (as $t_i$ is $i$'s first exceptional time), and $|E_{s-1}|\le M$; so $N_i\ge m$ and
$i\in B_1$. Thus $E_\tau\subseteq B_0\cup B_1$, contradicting $|E_\tau|>M$ on $\{|B_0|\le M/2,|B_1|\le M/2\}$. Hence
\begin{equation}\label{eq:Esparse}
  \Pr(|E_T|>nd^{-5})\le e^{-c\,nd^{-12}}+(\log n)^{-\omega(1)}.
\end{equation}

\emph{Level recursion.} Work on $\mathcal G=\{|E_T|\le nd^{-5}\}$. For $r\ge0$ set the strict level sets
$\mathcal M_r(t)=\{i:\max_{s\le t}\rangeop_i(s)>65\Hscale+3r\}$; since ranges are integers, climbing from level $r-1$ to $r$
forces three genuine upward steps above the previous threshold. \emph{Base.} If $i\in\mathcal M_0(T)$, then between its last
visit below $64\Hscale$ and first exceeding $65\Hscale$ there are $\ge m_0:=\lfloor\Hscale/2\rfloor$ arrivals at which
$\rangeop_i$ increases with pre-arrival range $\ge 64\Hscale$, so $i\in E_{t-1}$ and (R3) supplies a co-present helper. On
$\mathcal G$, \eqref{eq:incidence} with $|Q|\le nd^{-5}$ gives $\Pr(i\in\mathcal M_0(T),\mathcal G)\le\binom{T}{m_0}(2d^2\cdot nd^{-5}/n^2)^{m_0}\le(C/(m_0 d^3))^{m_0}$,
so with $a_0=(\log n)^{-2}$, $\Pr(|\mathcal M_0(T)|>na_0,\mathcal G)\le(\log n)^2(C/(m_0d^3))^{m_0}=(\log n)^{-\omega(1)}$.
\emph{Induction.} Put $a_r=(\log n)^{-2\cdot2^r}$, so $a_r=a_{r-1}^2$. If $i\in\mathcal M_r(T)$ then it has three increase-times
$t_u$ with pre-arrival range $>65\Hscale+3(r-1)$, each (by (R3)) with a helper $j_u\in S_{t_u}\setminus\{i\}$ of range
$>65\Hscale+3(r-1)$, i.e.\ $j_u\in\mathcal M_{r-1}(t_u-1)$. On $\{|\mathcal M_{r-1}(T)|\le na_{r-1}\}$, a union over ordered
time-triples with \eqref{eq:incidence} ($|Q|\le na_{r-1}$) gives $\Pr(i\in\mathcal M_r(T),\,|\mathcal M_{r-1}(T)|\le na_{r-1})\le C d^6 a_{r-1}^3$,
so by Markov $\Pr(|\mathcal M_r(T)|>na_r,\,|\mathcal M_{r-1}(T)|\le na_{r-1})\le Cd^6 a_{r-1}^3/a_r=Cd^6 a_{r-1}$. Summing for
$1\le r\le r_*:=\lceil 2\Hscale\rceil$ gives $\sum_r Cd^6 a_{r-1}\le CHd^6/(\log n)^2=o(1)$. Finally
$na_{r_*}=\exp(\log n-2^{r_*+1}\Hscale)<1$, so $\mathcal M_{r_*}(T)=\varnothing$ \whp; as $65\Hscale+3r_*\le 71\Hscale+3<80\Hscale$,
$\max_{t,i}\rangeop_i<80\Hscale$. Combining \eqref{eq:Esparse} with the level bounds proves the lemma.
\end{proof}

\paragraph{Our algorithm meets the interface.}
The minimum-load repair satisfies (R1)--(R3). \emph{(R1):} the repaired $i^*$ is exceptional, so $\rangeop_{i^*}(t-1)\ge 64\Hscale\ge1$;
incrementing a minimum-load color makes the new load $\min_c+1\le\max_c$, so the maximum is unchanged and the minimum can only
rise, hence the range does not increase. \emph{(R2):} the color differs from the seed only when the repair fires, which
requires $i^*\in S_t\cap E_{t-1}$, so $S_t\cap E_{t-1}\ne\varnothing$. \emph{(R3):} if an exceptional $i\in S_t\cap E_{t-1}$
has its range increase, it was not repaired (the repair is range-non-increasing by (R1)), so $i$ is not the unique strict
range-maximum of $S_t$; hence some $j\in S_t\setminus\{i\}$ has $\rangeop_j(t-1)\ge\rangeop_i(t-1)$. Thus \Cref{lem:transfer}
applies and $\max_{t,i}\rangeop_i<80\Hscale$ \whp, completing \Cref{lem:online-upper}. \qed

\begin{remark}
Condition (R1) is not used in the proof of \Cref{lem:transfer} beyond verifying (R3) for the concrete rule; (R2) drives the
exceptional-set sparsity and (R3) drives the level recursion. Nothing in the proof inspects which of the $k$ colors is used:
the count $D_i$, the contact count $N_i$, and the helper condition are all about which arrivals touch which coordinates, so
the bound holds for every $k\ge2$ for which the seed tail \eqref{eq:seed-tail} holds, with no dependence on $k$. Since the
seed tail holds for every $k\le d$ and every fixed $k$, so does \Cref{lem:online-upper}.
\end{remark}
