% appD-bc.tex -- the broadening-cost lemma + the phase-conveyor obstruction (for Section 6).
\section{The broadening-cost lemma and the phase conveyor}\label{app:bc}

This appendix supports the open problem of \Cref{sec:open}, and nothing in it feeds back into the main results: its
statements are not used by \Cref{thm:offline,thm:online,thm:separation}. The two objects we develop here are the adaptive
broadening-cost estimate, which is the positive input to the symmetric approach, and the deterministic phase conveyor, which
is the obstruction that rules out the expectation-level version of the asymmetric approach. Together they explain precisely
why the bracket of \Cref{sec:open} is hard to close from either side. We prove \Cref{lem:bc} in full, since it is a clean
probabilistic estimate; the phase conveyor is an informal deterministic construction whose verification we only sketch,
because it serves only to delimit the open problem and not to support any theorem.

\paragraph{The rank-greedy rule and its objects.}
The symmetric route studies the \emph{rank-greedy rule}: on each arrival $t$ the online algorithm assigns the seed color to
every coordinate except possibly one, which it repairs by the color minimizing that coordinate's rank $\rank$ of
\eqref{eq:rank} (the symmetric counterpart of the offline color-tree, breaking ties arbitrarily). For a coordinate $i$ at
rank $q$, its \emph{rank-$q$ forbidden set} is the set of colors whose use would not lower the rank below $q$ (the colors at
or near the maximum load); a color outside the forbidden set is one the rule would prefer. A \emph{broadening} of $i$ at
range $r$ is a step that increments a color one below the maximum, enlarging the argmax set and raising the rank by one
without changing the range. The rule's \emph{cover certificate} is the bookkeeping that, whenever the rule is forced to
broaden $i$ instead of using a cheaper out-of-forbidden-set color, names a distinct co-present coordinate $j$ in the support
whose rank is at least as high as the rank that cheaper color would have produced; this is what charges each broadening to a
coincidence of two high-range coordinates. These objects are internal to the symmetric analysis and are used only in
\Cref{lem:bc} below.

\paragraph{The broadening-cost lemma.}
We first set up the stopped process the estimate lives on. Fix a range level $r\ge2$ and a density $\eta\in(0,1)$. Let
$E_r(t)=\{i:\max_{u\le t}\rangeop_i(u)\ge r\}$ be the set of coordinates that have reached range at least $r$ by time $t$;
this set only grows in $t$. Stop the process the first time this set exceeds an $\eta$ fraction of all coordinates, at
$\sigma_{r,\eta}=\inf\{t:|E_r(t)|>\eta n\}$, so that up to the stopping time we always have $|E_r(t)|\le\eta n$. Let $B_{i,r}$
count the within-level broadenings of coordinate $i$ at range $r$, that is, the number of times $i$'s argmax set grows while
$i$ sits at range $r$, counted before the horizon $T\wedge\sigma_{r,\eta}$ and before $i$ first climbs to range $r+1$. The
lemma bounds the falling factorial moments $(B_{i,r})_a=B_{i,r}(B_{i,r}-1)\cdots(B_{i,r}-a+1)$, summed over coordinates, which
is the form a moment hierarchy consumes.

\begin{lemma}[Broadening cost]\label{lem:bc}
For every $r\ge2$ and integer $a\ge1$,
$\E\sum_{i=1}^n (B_{i,r})_a\le n\eta\,(2\tau_+ d^2\eta)^{a}$, uniformly over adaptive stopping-time epochs.
\end{lemma}

\begin{proof}
The argument is a charging argument followed by a sequential second-moment count. The point of the charging step is to show
that a broadening of $i$ can only happen when some \emph{other} coordinate that has already reached range $r$ is present in
the same support, so that broadenings are rare events tied to coincidences of two high-range coordinates.

Consider a broadening of $i$ at range $r$. By the definition of the rank, such a broadening reaches rank
$q=(k-1)(r-1)+\maxmult+1$. Because $r\ge2$, the coordinate $i$ has at least one color whose load is at most $\max-2$, and
this color lies outside the rank-$q$ forbidden set, so the rank-greedy rule would otherwise have used it. The cover
certificate of the rank-greedy rule then forces some other supported coordinate $j\ne i$ to cover $i$ at this step: the
broadening can only be the chosen move if a competing high-rank coordinate is present to block the cheaper alternatives.
Concretely, the rule moves $i$ to its broadening color only when every color of rank below $q$ on $i$ is already claimed by a
co-present coordinate that ranks at least as high, and the rank-cover certificate names, for each such blocked color, a
distinct witness $j$ in the support whose own rank meets or exceeds the rank that color would otherwise have produced. We read
off the range of any such witness. Any coordinate of range at most $r-1$ has post-move rank at most
$(k-1)(r-1)+1$, since its range factor contributes at most $(k-1)(r-1)$ and its multiplicity factor at most $1$ before the
move; this is strictly less than $q=(k-1)(r-1)+\maxmult+1$ for $\maxmult\ge1$. So a witness of range below $r$ cannot block a
rank-$q$ color, and the covering coordinate $j$ must already lie in $E_r(t-1)$. Hence both $i$ and a distinct $j$ lie in
$S_t\cap E_r(t-1)$ at the time of the broadening, and the witness $j$ is determined by the certificate, so distinct
broadenings of $i$ within the same arrival are charged to distinct witnesses. This is the charge: every broadening of $i$ is
charged to the simultaneous presence of $i$ and some already-range-$r$ coordinate $j$ in the support.

It follows that $B_{i,r}$ is dominated by the count $Y_i$ of arrival steps at which $i\in S_t$ and
$(S_t\setminus\{i\})\cap E_r(t-1)\ne\varnothing$, that is, steps where $i$ and at least one other range-$r$ coordinate share
the support. We now bound the factorial moments of $Y_i$. At any fixed step, the support $S_t$ is a uniform $d$-subset, so
the probability that $i\in S_t$ is $d/n$, and conditionally on $i\in S_t$ the remaining $d-1$ slots are a uniform
$(d-1)$-subset of the other $n-1$ coordinates; the probability that these slots meet the set $E_r(t-1)$ of size
$|E_r|\le\eta n$ is, by a union bound over the $d-1\le d$ slots, at most $(d-1)\,|E_r|/(n-1)\le d|E_r|/n$. The per-step
conditional probability of a charged step is therefore at most $(d/n)\cdot(d|E_r|/n)\le 2d^2\eta/n=:p$, where the factor $2$
absorbs the $n/(n-1)$ correction. To bound $\E\sum_i(Y_i)_a$ we expand the falling factorial
$(Y_i)_a=\sum y$ as a sum over ordered $a$-tuples of \emph{distinct} contributing steps for coordinate $i$, so that
$\sum_i(Y_i)_a$ counts ordered $a$-tuples of distinct steps each charged to a common coordinate $i$. We condition sequentially
on the steps in increasing time order. At the first of the $a$ steps the anchoring coordinate $i$ must lie in
$E_r(t-1)\subseteq E_r$, so there are at most $|E_r|\le\eta n$ choices for it; each of the $a$ steps then contributes the
per-step charge probability at most $p$ for the pair $(i,j)$ to be co-present. Summing the $\binom T a$ choices of unordered
step-set with the $a!$ orderings gives
\[
  \E\sum_i(Y_i)_a\ \le\ a!\binom T a\,n\eta\,p^{a}\ \le\ T^a\,n\eta\,p^{a}\ =\ n\eta\,(Tp)^{a}\ \le\ n\eta\,(2\tau_+ d^2\eta)^{a},
\]
where the middle step uses $a!\binom T a\le T^a$ and the last step uses $T\le\tau_+ n$ to write
$Tp\le T\cdot 2d^2\eta/n\le 2\tau_+ d^2\eta$. Because the supports at future steps remain i.i.d.\ uniform $d$-sets
regardless of the past, the per-step bound $p$ does not depend on the history, so the whole estimate is uniform over the
adaptive stopping-time epoch, as claimed.
\end{proof}

It is worth seeing how this clean estimate nonetheless produces the $\log k$ gap. Combined with the level-raising cover
bound, which says a coordinate reaches range $r+1$ only after accumulating enough range-$r$ broadenings, \Cref{lem:bc} feeds
into a symmetric range recurrence. The factorial-moment bound $n\eta(2\tau_+d^2\eta)^a$ behaves like the moments of a Poisson
variable of mean $\Th(d^2\eta)$, and demanding $k$ such broadenings to force a raise turns this into
$\eta_{r+1}\lesssim(C\eta_r)^k$ for the fraction $\eta_r$ at range $r$. Iterating gives $\eta\le e^{-ck^{j}}$ after $j$
levels, so making this beat a union bound over the $n$ coordinates requires $k^{j}\gtrsim\log n$, that is
$j\gtrsim\loglog n/\log k$, while the bracket only allows $j$ up to $\loglog n/k$. The discrepancy is the $\log k$ gap of
\Cref{sec:open}. As explained there, the gap is intrinsic to any permutation-invariant statistic, since the skip
configuration forces the full $k^{r}$ branching that the symmetric recurrence records; the estimate above is tight, and the
loss is in the symmetry of the statistic and not in the bound.

\paragraph{The phase conveyor.}
We turn to the asymmetric route and the construction that defeats its expectation-level form. The asymmetric rule we analyze
is \emph{LeftGap}, the color-sensitive analogue of Always-Go-Left. For a coordinate $i$ and color $c$ write
$g_{i,c}=x_{i,c}-\min_b x_{i,b}$ for the gap of color $c$ above the minimum load on $i$, and let $h_c=\max_{i\in S_t}g_{i,c}$
be the largest such gap over the current support. LeftGap picks the color $c$ that minimizes the pair $(h_c,c)$
lexicographically, so it prefers colors with small maximum gap and breaks ties toward the lowest color index, the
``leftmost'' direction. To track how gaps propagate we encode the event $g_{i,c}\ge\ell$ as a token of phase
$k(\ell-1)+c$, so that phases increase first through the color index $c$ within a fixed gap level $\ell$ and then jump to the
next level. A token of phase $u$ is \emph{principal} if it is the first coordinate to reach phase $u$ along the run, so that
principal tokens are the ones a moment count over phases would track. In a typical instance the number of coordinates
reaching a high phase obeys the staggered-Fibonacci recurrence $F_k$, and the conjectured bound would follow from an
expectation-level tail on the number of high-phase tokens. The following construction shows, at the level of a sketch, that
no such expectation-level tail can hold; we present it as an informal obstruction, not a proven proposition, since its
verification is by inspection of one explicit schedule.

\paragraph{Construction (phase conveyor; informal).}
For $k=d\ge3$ and every $q\ge1$ there is a deterministic sequence of $q$ exactly-$k$-sparse supports on which LeftGap creates
a principal token of each phase $1,2,\dots,q$ on the successive arrivals, using $\Oh(kq)$ coordinates. Consequently, if
$Z_u$ counts principal LeftGap tokens of phase $\ge u$, then $\E Z_u\ge\binom nk^{-\lceil u\rceil}$, which for
$k=\lfloor\Hscale^2/(2\log\Hscale)\rfloor$ exceeds $e^{-cF_k(m)}$ at $u=kB+m$ for any fixed $m>\Hscale/\log2$. So no bound
$\E Z_{kB+m}\le Cnd\,e^{-cF_k(m)}$ can hold on a window reaching $m=\Th(\Hscale)$.

The mechanism behind the proposition is a cycle invariant that lets a fixed set of coordinates be reused indefinitely. The
conveyor keeps three core coordinates alive, an anchor $a$ and two carriers $b_1$ and $b_j$, and feeds in fresh filler
coordinates each cycle. After $r$ completed cycles it maintains the gap profiles
\[
  g_a=(0,r,\dots,r),\qquad g_{b_1}=(r,0,\dots,0),\qquad g_{b_j}=r(e_1+e_j),
\]
so the anchor carries a high gap on every color except the first, carrier $b_1$ carries a high gap on the first color alone,
and carrier $b_j$ carries a high gap on the two colors $1$ and $j$. These profiles are arranged so that LeftGap, following its
leftmost tie-breaking, is forced to advance exactly one of the carriers each cycle while restoring the others. We trace one
cycle. When a fresh arrival presents the core together with $\Oh(k)$ fillers, the value $(h_c,c)$ that LeftGap minimizes is
realized at the unique color where the three core profiles leave the smallest maximum gap; the leftmost tie-break sends the
move to the first color that is not already saturated on the anchor, which by the shape of $g_a,g_{b_1},g_{b_j}$ is always the
designated carrier color of that cycle. The move raises that carrier's gap by one on its color and leaves the other two cores
restorable by the fillers, which soak up the off-color increments and are then discarded. Cycling through the colors
$c=1,\dots,k$ in turn advances the carriers one color at a time, and one full pass of $k$ such cycles advances the maximum
phase by exactly $k$, that is, by one full gap level. The advance is therefore linear in the number of cycles and not
Fibonacci: the conveyor manufactures one full level of $k$ phases per cycle-pass using only the recycled cores plus $\Oh(k)$
fillers per cycle, so $\Oh(kq)$ coordinates suffice to reach phase $q$, instead of the exponentially many coordinates a
typical instance would spend, where the high-phase count grows only as the staggered-Fibonacci $F_k$. This is exactly why it
inflates the expectation past the $e^{-cF_k(m)}$ that the asymmetric approach needs.

The construction is, however, vanishingly rare. Each of the $q$ arrivals must present exactly the prescribed
exactly-$k$-sparse support, and a fixed $k$-subset is the realized support of one uniform arrival with probability
$\binom nk^{-1}$; since the $q$ arrivals are independent, realizing the whole prescribed sequence has probability
\[
  \binom nk^{-q}\ =\ \exp\!\big(-q\log\tbinom nk\big)\ =\ \exp(-(\log n)\,\Hscale^{\Oh(1)}),
\]
using $\log\binom nk=\Th(k\log n)$ and $q,k=\Hscale^{\Oh(1)}$, which is smaller than $n^{-A}$ for every fixed $A$. So the
conveyor does not contradict any high-probability bound, and it is not an online lower bound; what it refutes is the
expectation-level tail, because the expectation $\E Z_u$ sums over all instances and so collects the rare conveyor's huge
contribution undamped, and with it the stopping-time-uniform mixed-profile factorization that an expectation argument would
rely on. This pins down what a correct argument must look like. The corrected, sufficient statement for \Cref{prob:growing}
is the truncated bound $\Pr(Z_{kB+m}>0)\le Cnd\,e^{-cF_k(m)}+n^{-A}$, which adds an explicit $n^{-A}$ slack to absorb the rare
conveyor and so is consistent with both the conveyor and the conjecture. Proving it would require a cycle-aware
``fold-or-Fibonacci'' count: an accounting that, every time a coordinate is reused, charges the reuse to enough genuinely
independent earlier births that the recycled conveyor cannot be assembled cheaply. Producing such a count is exactly the
content of the open problem, and it remains open.
