% sec3-offline.tex -- The offline characterization (technical centerpiece). Full proofs -> appB.
\section{The offline characterization}\label{sec:offline}

This section proves the offline value exactly.

\begin{theorem}[Offline characterization]\label{thm:offline}
There is an absolute constant $C_0$ such that for $k\le d$, $d,k=(\log n)^{o(1)}$, $k\ge C_0\log^5(dk)$, and $\lam\ge1$,
\[
  \offdisc_k\ =\ \Th\big(\Psifn(\lam,\log k)\big)\qquad\whp,
\]
with $\lam=Td/(nk)$ and $\Psifn$ as in \eqref{eq:Psi}. (Here $\offdisc_k$ is the final discrepancy, \Cref{sec:model}; the
occupancy condition $\lam\ge1$ is equivalent to $k\le Td/n$ and holds for all $k\le d$ once $T\ge n$.)
\end{theorem}

The hypothesis $k\ge C_0\log^5(dk)$ is what lets the per-coordinate spread beat the union over the $k^T$ colorings; it holds
for all but the smallest $k$ in the range, and excludes $k=2$, consistent with the two-color value being $\Th(\sqrt d)$
and not $\Th(\Psifn)$. The role of the hypothesis is worth stating plainly. The lower bound is a first-moment argument over
the $k^T$ offline colorings, and a coloring survives only if its per-coordinate spread is atypically small. The spread is a
sum of $n$ nearly independent per-coordinate contributions, so its lower tail is exponentially small in $n$, while the union
runs over $e^{\Th(n\log k)}$ colorings. The fifth power of $\log(dk)$ is the slack we need for the per-coordinate
concentration to dominate this entropy with room to spare; once $k$ is below that threshold the entropy can win and the
$\Psifn$ prediction need not hold, which is exactly what happens at $k=2$.

The matching upper and lower bounds are \Cref{lem:offline-lower,lem:offline-upper}. The value
$\Psifn(\lam,\log k)$ is the spread of $k$ independent $\Pois(\lam)$ loads at a coordinate: the lower bound says a
coordinate cannot avoid this spread, and the upper bound exhibits a coloring that matches it at every coordinate at once.
The two halves are different kinds of statement. The lower bound is about \emph{every} coloring and is proved by counting;
the upper bound is about \emph{one} coloring and is proved by construction. They meet at $\Psifn$ because the same Poisson
extremal problem sets both the floor that no coloring can dig under and the ceiling that a careful recursive coloring can
reach. For fixed $d$, $k=2$ recovers the classical $\Th(\sqrt d)$ value of random $d$-regular set systems
\cite{BansalMeka2019,EzraLovett2016,HobergRothvoss2019}; the content here is the dependence on $k$.

\subsection{The occupancy lower bound}\label{sec:offline-lower}

\begin{lemma}[Lower bound]\label{lem:offline-lower}
Under the hypotheses of \Cref{thm:offline}, $\offdisc_k\ge c\,\Psifn(\lam,\log k)$ \whp, for an absolute constant $c>0$.
\end{lemma}

\paragraph{Idea.}
The lower bound is a first-moment, or union-bound, argument: we show that the expected number of offline colorings whose
every coordinate has range below $c\Psifn$ is $o(1)$, so \whp\ no such coloring exists. Three moves make this tractable. The
first move replaces the conditioning. Working directly with exactly-$d$ columns couples the coordinates, so instead we draw
each entry as an independent $\Bin(1,d/n)$ and condition on each column having exactly $d$ ones, which recovers $\mu_{d,n}$
exactly. The conditioning costs a factor $(\Oh(\sqrt d))^{T}$, which is $e^{\Oh(n\log d)}$ and is comfortably dominated by
the entropy $\Th(n\log k)$ of the union over the at most $k^{T}=e^{\Th(n\log k)}$ colorings; this is the first place the
hypothesis on $k$ pays for itself. The second move is that in the Bernoulli model the $k$ color-loads at a fixed coordinate
are now independent: color $c$'s load is $\Bin(m_c,d/n)$ with mean $\mu_c=m_c d/n$ and $\sum_c\mu_c=k\lam$, and the relevant
Poisson picture is in force.

Fix an offline coloring $\chi$ and suppose every coordinate has range at most $D$. The third move shows the loads spread no
matter how the coloring tries to balance them, through a dichotomy. Either some color carries mean $\gg\lam$, in which case
that color's load alone overshoots the rest by $\Psifn$ and the range is already large; or no color is heavy, in which case
a constant fraction of the colors have mean $\Th(\lam)$ and we are looking at the extreme value of many comparable Poisson
loads, which spans $\Psifn$. The Poisson calibration is the same one that defines $\Psifn$: by \eqref{eq:poisson-tail}, the
deviation at which the expected number of colors exceeding $\mu_c+t$ drops below one solves rate $=\log k$, giving
$t\asymp\sqrt{\lam\log k}+\log k/\log(e+\log k/\lam)=\Th(\Psifn)$. At a single coordinate the range therefore drops below
$\tfrac14\Psifn$ only if the $k$ loads conspire to stay in a narrow window, an event of probability at most $e^{-k^{\Om(1)}}$
(the occupancy-spread lower tail, \Cref{lem:occ-spread}, which requires $\lam\ge1$). To go from one coordinate to all of them
we use that the per-coordinate events are independent across the $n$ rows of the Poissonized matrix, so a fixed coloring
keeps every range small with probability at most $e^{-n\,k^{\Om(1)}}$. This beats the $e^{\Th(n\log k)}$ union over the
$k^T$ colorings precisely because $k\ge C_0\log^5(dk)$ makes $k^{\Om(1)}$ dominate $\log(dk)$. The full argument, with the
heavy/light dichotomy and the three regimes \eqref{eq:Psi-regimes}, is in \Cref{app:offline-lower-full}.

\subsection{Overview of the upper bound}\label{sec:offline-overview}

The upper bound is the technical heart. We must produce, from a single draw of the incidence matrix $A\in\{0,1\}^{n\times T}$,
one offline $k$-coloring whose every coordinate range is $\Oh(\Psifn)$. Three ingredients combine.

First, a \emph{structural lemma} (\Cref{lem:structural}) certifies that $A$ is simultaneously well behaved for
\emph{every} column subset $S\subseteq[T]$: the row counts $r_i(S)=\sum_{t\in S}A_{it}$ have light tails uniformly over
$S$. This uniformity is what lets an adaptive recursion choose its column subsets from the realized matrix without a fresh
appeal to randomness.

Second, a density-sensitive \emph{splitting lemma} (\Cref{lem:split}) shows that any column class $S$ can be partitioned
into two parts of any prescribed sizes so that each row's count splits in the same proportion up to an additive
$\Oh(\Psifn(\mu_s,\ell_s))$ error, where $\mu_s,\ell_s$ are the density parameters of $S$. The proof iterates the
Lovett--Meka lemma with thresholds tuned to the local density; an all-ones constraint of threshold $0$ keeps the part sizes
exact (\Cref{lem:LM}).

Third, a \emph{recursive color-tree} (\Cref{lem:offline-upper}) builds the $k$ classes by repeatedly splitting along a
balanced binary tree of depth $\log_2 k$. An error introduced at a node with $q$ leaves below it is attenuated by a factor
$\Oh(1/q)$ on the way to a single leaf, and the geometric decrease of $q$ down a root-to-leaf path makes the accumulated
error a convergent sum that reassembles exactly into $\Psifn(\lam,\log k)$, with no extra factor of $\log d$, $\sqrt{\log d}$,
or $\loglog n$.

\subsection{The structural lemma}\label{sec:structural}

The recursion below is adaptive: at each node it looks at the realized matrix and chooses which columns to split off, so the
column subsets it works with are not fixed in advance but read out of the same random draw it is trying to control. If we
wanted a tail bound for each subset on demand we would have to spend fresh randomness every time, and there is none left.
The structural lemma removes this difficulty by paying once, up front, for \emph{all} subsets at once. It says that on a
single high-probability event the row counts $r_i(S)$ are tame uniformly over every column set $S$, so the recursion may
treat the matrix as deterministically well behaved and never appeal to randomness again. The two displayed bounds are the
dense and sparse faces of the same control: \eqref{eq:struct-dense} bounds an exponential moment of the row counts when $S$
is dense enough that loads concentrate, and \eqref{eq:struct-sparse} bounds the number of overloaded rows when $S$ is sparse
and the right notion is a hard cap at $\Th(\Psifn(\mu_s,\ell_s))$. The threshold $\ell_s=\log(eT/s)$ is the per-subset entropy
budget: it is the log of how many subsets of size $s$ there are, normalized per column, and it is what the union bound must
overcome.

For $S\subseteq[T]$ with $s=|S|$ write $\mu_s=ds/n$ and $\ell_s=\log(eT/s)\ge 1$.

\begin{lemma}[All-subsets structural lemma]\label{lem:structural}
There is an absolute constant $K$, depending only on the bounds on $T/n$, such that \whp\ the following hold
simultaneously for all $S\subseteq[T]$.
\begin{align}
  \mu_s\ge\ell_s:&\qquad \sum_{i:\,r_i(S)>0}\exp\!\Big(-\tfrac{K^2\mu_s\ell_s}{16\,r_i(S)}\Big)\ \le\ \tfrac{s}{64};
    \label{eq:struct-dense}\\
  \mu_s<\ell_s:&\qquad \big|\{\,i:\ r_i(S)>K\,\Psifn(\mu_s,\ell_s)\,\}\big|\ \le\ \tfrac{s}{64}.
    \label{eq:struct-sparse}
\end{align}
\end{lemma}

The proof (\Cref{app:structural-full}) fixes $S$, bounds the relevant sum or count in expectation by a binomial Chernoff
estimate \eqref{eq:struct-dense} or a Bennett estimate \eqref{eq:struct-sparse} of order $e^{-B\ell_s}$ for a large
constant $B$, and then converts to a high-probability statement using that $\{r_i(S)\}_i$ is negatively associated
(\Cref{sec:NA}); the second summand of $\Psifn$ is exactly calibrated so that the Bennett exponent reaches $B\ell_s$. The
union over the at most $\binom{T}{s}\le e^{s\ell_s}$ subsets of each size converges because the per-subset failure
probability $e^{-cBs\ell_s}$ and the subset count $e^{s\ell_s}$ scale with the same $s\ell_s$, so a large absolute $B$ wins.
This single uniform event is what supports the adaptive recursion below.

\subsection{The splitting lemma}\label{sec:split}

The splitting lemma is the single operation the color-tree performs at every node: take a column class $S$ and cut it into
two pieces of prescribed sizes so that each row is divided in the same ratio, up to a small error. If we could split every
row perfectly in proportion there would be no discrepancy at all; the content is that we can come within
$\Oh(\Psifn(\mu_s,\ell_s))$ of perfect on every row simultaneously while keeping the two part sizes exactly as ordered. Two
features make this the right tool. It is \emph{density-sensitive}: the error scales with $\Psifn$ at the local density of
$S$, not at the global density, which is what lets the per-node errors shrink as the tree descends into sparser subsets. And
it is \emph{exact-cardinality}: the part sizes are hit on the nose, so the leaves end up with the target color sizes $m_c$
and the only thing left to control is the row counts. The proof is a Lovett--Meka rounding driven by the structural lemma:
the dense bound \eqref{eq:struct-dense} is exactly the budget condition $\sum_j e^{-c_j^2/16}\le m/16$ that
\Cref{lem:LM} needs, so the structural event is what makes the rounding legal.

\begin{lemma}[Density-sensitive exact-cardinality splitting]\label{lem:split}
On the event of \Cref{lem:structural}, for every $S\subseteq[T]$ with $s=|S|$ and every integer $0\le a\le s$ there is
$Y\subseteq S$ with $|Y|=a$ and
\[
  \max_{i}\ \Big|\,r_i(Y)-\tfrac{a}{s}\,r_i(S)\,\Big|\ \le\ C\,\Psifn(\mu_s,\ell_s),
\]
for an absolute constant $C$.
\end{lemma}

The proof (\Cref{app:split-full}) runs \Cref{lem:LM} from the fractional point $x^{(0)}_t=2a/s-1$, always including the
all-ones vector with threshold $0$ so that $\sum_t\sigma_t$ stays fixed and $|Y|=a$ exactly. While the alive set is dense
($\mu_A\ge\ell_A$) each row $i$ with $r_i(A)>0$ gets threshold $c_i=K\sqrt{\mu_A\ell_A}/\sqrt{r_i(A)}$; by
\eqref{eq:struct-dense} the threshold budget closes and each iteration moves a row by at most $K\sqrt{\mu_A\ell_A}$. When the
alive set turns sparse ($\mu_A<\ell_A$) every row above $K\Psifn(\mu_A,\ell_A)$ is constrained and every row below it is
released for good, contributing at most $2K\Psifn(\mu_A,\ell_A)$ thereafter. Because the alive set halves each iteration,
the per-stage errors form a convergent geometric sum, bounded in \Cref{app:split-full} by $C\Psifn(\mu_s,\ell_s)$.

\subsection{The recursive color-tree}\label{sec:tree}

The color-tree assembles the splitting lemma into a full $k$-coloring and is where the answer $\Psifn(\lam,\log k)$ is
finally produced. The naive way to build $k$ classes by splitting would apply the worst-case error at every step and pile up
a factor of $\log k$ or worse; the design here avoids that. Colors are the leaves of a balanced binary tree of depth
$\log_2 k$, and a class is built by recursively splitting the column set at each internal node into its two subtrees. The
key accounting fact is attenuation: an error committed at a node with $q$ leaves below it is a discrepancy spread over a
class of size $\Th(qT/k)$, so by the time the recursion has narrowed down to one leaf that error has been divided by
$\Th(q)$. Because the densities along a root-to-leaf path are geometrically spaced and the $q$'s halve at each step, the
attenuated per-node errors form a convergent series and not a growing sum, and the series totals exactly $\Psifn(\lam,\log k)$
with no stray factor of $\log d$, $\sqrt{\log d}$, or $\loglog n$. This cancellation is the reason the offline value collapses
to the clean Poisson spread, and it is the technical heart of the whole section.

\begin{lemma}[Upper bound]\label{lem:offline-upper}
On the event of \Cref{lem:structural}, there is an offline $k$-coloring with final discrepancy
$\max_{i}\rangeop_i(T)\le C\,\Psifn(\lam,\log k)$.
\end{lemma}

\paragraph{Idea.}
The plan is to reduce the range at each coordinate to a sum of attenuated splitting errors and then sum that series.
Fix target class sizes $m_c\in\{\lfloor T/k\rfloor,\lceil T/k\rceil\}$ summing to $T$, and a balanced binary tree whose $k$
leaves are the colors. A node $v$ with $q_v$ leaves below it holds a column set of size $M_v=\sum_{c\text{ below }v}m_c$;
split it into its two children by \Cref{lem:split} with the exact child sizes. Since the offline discrepancy is the range of
the final loads, it suffices to bound each leaf's final deviation $|x_{i,c}(T)-(m_c/T)\deg(i)|$, where $\deg(i)$ is
coordinate $i$'s total count. Unrolling the splits along a root-to-leaf path, the error injected at a $q$-leaf node reaches
the leaf scaled by $\Oh(1/q)$, so
\[
  \Big|x_{i,c}(T)-\tfrac{m_c}{T}\deg(i)\Big|\ \le\ C\sum_{q\text{ on path}}\frac{\Psifn(\Th(\lam q),\,1+\log(k/q))}{q}.
\]
The $q$'s decrease geometrically, so this is a geometric-type series and each summand is dominated by its endpoints. The
square-root summand contributes
$\sqrt\lam\sum_q\sqrt{(1+\log(k/q))/q}=\Oh(\sqrt{\lam\log k})$, the Gaussian part of $\Psifn$, because the
$\sqrt{1/q}$ weights are summable and the largest term sits at the leaf end. The Bennett summand contributes
$\Oh(\sqrt{\lam\log k}+\log k/\log(e+\log k/\lam))$ after splitting the sum at $q=\sqrt{\log k/\lam}$, the depth at which the
local density crosses its own seam, so one side of the split feeds the square-root term and the other feeds the
bounded-occupancy term. The two pieces
reassemble into $\Psifn(\lam,\log k)$, and the ideal terms $(m_c/T)\deg(i)$ differ across colors by at most
$\deg(i)/T\le 1$, where $\deg(i)=\deg_T(i)$ is the final degree. This bounds the final range $\rangeop_i(T)$, which is
$\offdisc_k$ by definition. The full calculation is in \Cref{app:tree-full}.

\medskip
\Cref{thm:offline} follows by combining \Cref{lem:offline-lower,lem:offline-upper}.
