% sec4-online.tex -- Online fixed-k: P2 lower + multicolor A-T upper. Full proofs -> appC.
\section{The online value for fixed $k$}\label{sec:online}

Online, the picture is the opposite of \Cref{thm:offline}. In the offline setting extra colors give the algorithm more
room to balance and the discrepancy drops as $k$ grows; online, the colors must be committed one arrival at a time with no
view of the future, and this freedom evaporates. In the fixed-$k$ regime more colors do not help: the online value sits at
$\Th(\loglog n)$ for every constant $k$, the same order it takes at $k=2$. The reason, made precise below, is that the
adversary can fold the $k$ colors back down to two by a fixed signing and recover the two-color lower bound, while the
algorithm cannot use the extra colors to push the discrepancy below the two-color barrier.

\begin{theorem}[Online, fixed $k$]\label{thm:online}
For every fixed $k\ge 2$, $\ondisc_k=\Th(\loglog n)$ \whp, independent of $d$ in the range $2\le d\le\Hscale^2/\log\Hscale$.
\end{theorem}

The two bounds are proved separately and then combined. The lower bound is \Cref{lem:online-lower} ($\Om(\loglog n/k)$, for
all $k$); the upper bound is \Cref{lem:online-upper} ($\Oh(\loglog n)$, for $2\le k\le d$). Together they pin the online value
at $\ondisc_k=\Th(\loglog n)$ once $k$ is constant, with no dependence on the sparsity $d$ across the whole range
$2\le d\le\Hscale^2/\log\Hscale$. The two-color case $k=2$ is the theorem of Altschuler and Tikhomirov~\cite{AT2025}. Our
contribution is that both bounds extend across the palette range: the lower bound by a signing that projects $k$ colors onto
two, and the upper bound by a $k$-color seed-and-repair algorithm whose analysis reduces, color count and all, to the
two-color analysis of \cite{AT2025}. In particular the upper bound stays $\Oh(\loglog n)$ as $k$ grows up to $d$, so within
this range adding colors does not raise the online order.

\subsection{The lower bound}\label{sec:online-lower}

\begin{lemma}[Online lower bound]\label{lem:online-lower}
For all $k\ge 2$, $\ondisc_k\ge\Om(\loglog n/k)$.
\end{lemma}

\paragraph{Idea.}
The lower bound says that no online $k$-coloring can beat $\Om(\loglog n/k)$, so for constant $k$ the online value is
$\Om(\loglog n)$. The mechanism is to hide a two-color problem inside the $k$-color one. Project the $k$ colors onto two
groups by a fixed signing $\xi:[k]\to\{+1,-\rho\}$ and apply the two-color online lower bound \cite{AT2025} to the projected
process. For even $k$, split $[k]$ into equal halves with $\xi\equiv+1$ on one and $\xi\equiv-1$ on the other. The signed sum
$\sum_c\xi(c)x_{i,c}$ at any coordinate is a difference between a sum of $k/2$ loads and another sum of $k/2$ loads, and since
all $k$ loads lie within $\rangeop_i$ of one another, $|\sum_c\xi(c)x_{i,c}|\le (k/2)\,\rangeop_i$. Thus an online
$k$-coloring induces, through $\xi$, an online two-coloring whose discrepancy is at most $(k/2)\,\rangeop_i$. The two-color
bound $\Om(\loglog n)$ applies to this induced coloring and forces $(k/2)\,\rangeop_i\ge\Om(\loglog n)$, that is
$\rangeop_i\ge\Om(\loglog n/k)$. For odd $k=2a+1$ an equal split is impossible, so we use the asymmetric zero-sum weighting
$\xi\equiv 1$ on the size-$(a+1)$ part and $\xi\equiv -(a+1)/a$ on the size-$a$ part, which still has $\sum_c\xi(c)=0$. After
rescaling so that the positive step is $+1$, the action set is $\{+1,-\rho\}$ with $\rho=a/(a+1)\in[\tfrac12,1]$. The
two-color recurrence of \cite{AT2025} drives its lower bound through $\Th(\loglog n)$ stages, and each stage grows by a fixed
amount that does not depend on the exact step ratio as long as it stays in the band $[\tfrac12,1]$; the only effect of
$\rho\ne 1$ is a constant factor at most $2$ in the number of stages. So the same $\Om(\loglog n)$ bound follows for odd $k$
as well (\Cref{app:online-lower-full}).

\subsection{The upper bound}\label{sec:online-upper}

\begin{lemma}[Online upper bound]\label{lem:online-upper}
For every $k$ that is either at most $d$ or bounded by an absolute constant, there is an online $k$-coloring with
$\max_{t,i}\rangeop_i(t)\le 80\,\loglog n$ \whp\ (when $T=n$; for $T=\Th(n)$ the constant changes only). In particular this
covers every fixed $k\ge2$ across the whole range $2\le d\le\Hscale^2/\log\Hscale$, and every $k\le d$.
\end{lemma}

\paragraph{The algorithm.}
The upper bound is achieved by a seed-and-repair algorithm: color almost everything by a fixed random seed, and step in only
to repair the rare coordinate that has drifted dangerously high. Before any arrivals, pre-sample seed colors
$\widetilde\chi_t\sim\Unif[k]$, one per future arrival, independently; this is private randomness fixed in advance and used
only when arrival $t$ appears, so the algorithm stays online. These seeds would already give a good coloring on their own,
and the repair rule only intervenes to control the few coordinates whose range grows large under the seed. Track the
\emph{exceptional set} $E_t$, the set of coordinates whose range has reached $64\Hscale$ by time $t$; once a coordinate
enters $E$ it stays, so $E_t$ is nondecreasing. On arrival $t$ with support $S_t$, look for a single coordinate that is both
already exceptional and the clear leader of its support: if there is a \emph{unique} coordinate $i^*\in S_t\cap E_{t-1}$
whose range strictly exceeds that of every other $j\in S_t$, assign $i^*$ a color that attains the minimum load
$\min_c x_{i^*,c}(t-1)$ at that coordinate. In every other case (no exceptional coordinate in the support, or no unique
strict maximum) assign the seed color $\widetilde\chi_t$ and leave the seed coloring untouched. The point of the
minimum-load assignment is that, at a coordinate whose range is already positive, incrementing a least-loaded color cannot
raise the spread between the largest and smallest loads, so the repair never increases the range of the coordinate it
repairs, for any $k$: the repaired coordinate is exceptional, hence of range $\ge 64\Hscale\ge1$, so the new load
$\min_c+1$ is at most the old maximum, the maximum does not move, and the minimum can only rise. A coordinate can therefore cross to a higher range
level only on an arrival where it is \emph{not} the unique strict maximum of its support, which means some other
already-exceptional coordinate is present in the same arrival to absorb the blame.

\paragraph{Idea.}
The whole analysis turns on the observation just made: assigning a minimum-load color never raises that coordinate's range,
so a coordinate can only cross a range level when it is \emph{not} the unique strict maximum of its support, and that forces
another already-exceptional coordinate to be present in the same arrival. This is the one structural fact the rest of the
argument needs, and it holds for every $k$. The proof has three parts (\Cref{app:online-upper-full}). Part~1 controls the
seed alone: a maximal Bernstein bound shows the seed-only range exceeds $32\Hscale$ with probability at most $d^{-28}$. The
relevant martingale has predictable variance $\Th(d/k)$, and a short two-regime calculation (\Cref{app:online-upper-full})
checks that the $k^2$-pair union is dominated by the exponent, so the bound holds for every $k\le d$ and every fixed $k$.
Parts~2 and~3 are the category and level-set arguments of \cite{AT2025}, which we apply through an explicit interface: any
seed-and-repair rule for which (i) the repair does not raise the range of the coordinate it repairs, (ii) the color differs
from the seed only when the support holds a coordinate already at range $\ge 64\Hscale$, and (iii) an exceptional
coordinate's range increases only when the support holds another coordinate of at least its range, inherits the
$\Oh(\loglog n)$ bound. Property (iii) is the one the pure-seed coloring lacks, and our minimum-load repair supplies it. None
of (i)--(iii) mentions the number of colors, so the analysis applies for all $k\le d$. It shows the exceptional set has size at most
$n d^{-5}$. A double-exponential recurrence on the level sets $M(t,r)=\{i:\max_{s\le t}\rangeop_i(s)\ge 65\Hscale+3r\}$ then
collapses the count to zero by $r=2\Hscale$, giving $\max_{t,i}\rangeop_i<80\Hscale$ and hence $\Oh(\loglog n)$.

\medskip
\Cref{thm:online} follows by combining \Cref{lem:online-lower,lem:online-upper}.

\begin{remark}\label{rem:online-upper-allk}
\Cref{lem:online-upper} holds for every $k$ in the studied range $2\le k\le d$, so $\ondisc_k=\Oh(\loglog n)$ throughout. The
online upper bound is therefore $d$-independent, in contrast with the $d$-dependent offline threshold $\Psifn$ of
\Cref{thm:offline}. For growing $k$ the upper bound $\loglog n$ need not be tight; see \Cref{sec:separation,sec:open}.
\end{remark}
