% sec5-separation.tex -- The online-vs-offline separation + the growing-k bracket.
\section{The online--offline separation}\label{sec:separation}

The two halves of the analysis, the offline upper bound of \Cref{thm:offline} and the online lower bound of
\Cref{thm:online}, can now be placed side by side. Read together they say that, on the \emph{same} i.i.d.\ sparse input, the
online and offline values live on genuinely different scales, a multicolor analogue of the two-color separation of
\cite{AT2025}. In the two-color setting that separation already shows that a rule forced to commit to each column as it
arrives suffers an asymptotically larger imbalance than an offline colorer who sees the whole matrix; the statement below
records that the same phenomenon survives, and in fact sharpens, when the palette has any fixed number of colors.

\begin{theorem}[Separation, fixed $k$]\label{thm:separation}
For every fixed $k\ge 2$ and $2\le d\le\Hscale^2/\log\Hscale$,
\[
  \ondisc_k\ =\ \Th(\loglog n)\ \gg\ \offdisc_k\ =\ \Oh(\sqrt d)\ =\ o(\loglog n)\qquad\whp.
\]
\end{theorem}

\begin{proof}
There are three things to verify: that the online value sits at $\Th(\loglog n)$, that the offline value is $\Oh(\sqrt d)$,
and that the sparsity cap forces the second to be of strictly smaller order than the first.

The online value is $\Th(\loglog n)$ by \Cref{thm:online}, which supplies both directions at fixed $k$. The lower bound is
the heart of the matter: on the random i.i.d.\ stream of \Cref{def:model}, no online rule can keep the maximum per-coordinate
spread below order $\loglog n$, because a coloring committed column by column drives the rank statistic upward at the doubly
logarithmic rate forced by the self-similar branching of the allocation (\Cref{lem:online-lower}). The matching upper bound
(\Cref{lem:online-upper}) keeps the spread at $\Oh(\loglog n)$, so the online order is pinned at $\Th(\loglog n)$.

The offline value is $\Oh(\sqrt d)$. This needs only the trivial monotonicity $\offdisc_k\le\offdisc_2$: any $2$-coloring is
a $k$-coloring that leaves $k-2$ colors empty, so the optimal $k$-colorer does at least as well as the optimal $2$-colorer.
The two-color value is the classical discrepancy of a random sparse set system, $\offdisc_2=\Th(\sqrt d)$
\cite{BansalMeka2019,EzraLovett2016,HobergRothvoss2019}. Hence $\offdisc_k\le\offdisc_2=\Oh(\sqrt d)$ for every fixed $k$,
which is all the separation needs. (More colors in fact \emph{help}, and \Cref{thm:offline} pins the value at the smaller
$\Th(\Psifn(\lam,\log k))$ once $k$ grows past $C_0\log^5(dk)$; for fixed $k$ the $\log k$ argument is bounded and $\Psifn$ is
of order $\sqrt d$, consistent with the $\Oh(\sqrt d)$ here.)

It remains to compare the two scales. Under the sparsity cap $d\le\Hscale^2/\log\Hscale$ we may take square roots, and since
$d\mapsto\sqrt d$ is increasing this gives
\[
  \sqrt d\ \le\ \sqrt{\Hscale^2/\log\Hscale}\ =\ \frac{\Hscale}{\sqrt{\log\Hscale}} .
\]
Now $\Hscale=\loglog n\to\infty$ as $n\to\infty$, so $\log\Hscale\to\infty$ and the prefactor $1/\sqrt{\log\Hscale}\to0$.
Therefore $\sqrt d\le\Hscale/\sqrt{\log\Hscale}=o(\Hscale)$, that is $\sqrt d=o(\loglog n)$: the offline value is of strictly
smaller order than $\loglog n$, with the ratio $\offdisc_k/\loglog n=\Oh(1/\sqrt{\log\Hscale})\to0$ quantifying how much
smaller. Therefore $\offdisc_k=\Oh(\sqrt d)=o(\loglog n)=o(\ondisc_k)$, which is exactly the displayed chain of the theorem.
\end{proof}

A word on why the offline value here is quoted from the classical random-set-system bounds and not read off from
\Cref{thm:offline}. \Cref{thm:offline} is proved in the growing-palette regime $k\to\infty$, where its hypotheses are in
force, so it does not directly cover a fixed $k$. The two descriptions agree on the overlap of their ranges: the offline
formula $\Psifn(\lam,\log k)$ of \Cref{thm:offline} collapses to $\Th(\sqrt d)$ when $k$ is held constant, because the
$\log k$ argument is then a bounded constant and $\Psifn$ reduces to its Gaussian-spread regime. So the two routes give the
same offline order on the common domain, and we use whichever is licensed by its hypotheses.

For growing $k$ the online value is no longer pinned to a single power of $\loglog n$, and the clean separation above gives
way to a two-sided estimate. Combining \Cref{thm:offline} (through the trivial inequality $\ondisc_k\ge\offdisc_k$, since any
online coloring is in particular an offline one) with the online lower and upper bounds
\Cref{lem:online-lower,lem:online-upper} gives the following bracket.

\begin{corollary}[Growing-$k$ bracket]\label{cor:bracket}
For $C_0\log^5(dk)\le k\le d$ with $d,k=(\log n)^{o(1)}$ and $\lam\ge1$,
\[
  \max\!\Big\{\,\Psifn(\lam,\log k),\ \frac{\loglog n}{k}\,\Big\}\ \le\ \ondisc_k\ \le\ \loglog n\qquad\whp.
\]
(The floor $\loglog n/k$ alone, from \Cref{lem:online-lower}, needs no condition on $k$; the $\Psifn$ floor uses
\Cref{thm:offline} and hence its hypotheses $k\ge C_0\log^5(dk)$ and $\lam\ge1$.)
\end{corollary}

The two endpoints of this bracket can be far apart, and which of the two floors dominates depends on the size of $k$
relative to $\loglog n$. The lower bound is the maximum of two terms of very different origin. The term $\Psifn(\lam,\log k)$
is the offline floor: it is a price that even an all-seeing colorer must pay, inherited from \Cref{thm:offline}. The term
$\loglog n/k$ is online-specific: it is the cost the doubly logarithmic branching imposes on any column-by-column rule, and
it has no offline counterpart. For small $k$ the online-specific term $\loglog n/k$ is the larger of the two and strictly
exceeds the offline floor $\Psifn$, so online coloring is genuinely worse than offline coloring in that range, which is the
qualitative content of the separation extended past fixed $k$. As $k$ grows the online-specific term decays like $1/k$,
while the offline floor $\Psifn$ grows with its $\log k$ argument, so the two must cross.

The two floors cross. The online-specific floor $\loglog n/k$ decreases in $k$, while the occupancy floor
$\Psifn(\lam,\log k)$ increases in $k$ once the palette is large: in the bounded-occupancy corner $\lam=\Th(1)$ (that is
$k=\Th(d)$), \eqref{eq:Psi-regimes} gives $\Psifn=\Th(\log k/\loglog k)$, which grows, while at the other end small $k$ keeps
$\Psifn$ at $\Th(\sqrt d)$ and $\loglog n/k$ near $\loglog n$. So for small $k$ the online-specific floor $\loglog n/k$ is the
larger of the two and strictly exceeds $\Psifn$, the separation persisting past fixed $k$; for $k$ approaching $d$ the
occupancy floor $\Psifn$ takes over. The crossover happens at an intermediate palette where $k\,\Psifn(\lam,\log k)\asymp\loglog n$;
its exact location depends on $d$ through $\lam=\Th(d/k)$, and inside the resulting window the bracket leaves a genuine gap
between the lower and upper bounds. Pinning the exact growing-$k$ order in that gap is the subject of \Cref{sec:open}.

Two consequences are worth stating. First, the gap is of strictly larger order across the whole sparse range
$2\le d\le\Hscale^2/\log\Hscale$: the online value $\Th(\loglog n)$ exceeds the offline value $\Oh(\sqrt d)$ by a factor at
least $\Om(\sqrt{\log\Hscale})$, and by an unbounded factor up to $\Th(\loglog n)$ when $d=\Th(1)$. Second, the separation is
not an artifact of a single palette size: it persists for every fixed $k\ge2$, with the implied constants depending on $k$
only through bounded factors, so the online disadvantage is a stable structural feature of the problem and not a coincidence
at one value of $k$.
