% sec6-open.tex -- The growing-k frontier (open problem).
\section{The growing-$k$ regime}\label{sec:open}

\Cref{cor:bracket} leaves a gap between its lower and upper bounds for growing $k$. The purpose of this section is to say as
precisely as possible what would be needed to close that gap, and to explain why two natural lines of attack, one symmetric
and one asymmetric, each fall short for a concrete and identifiable reason. To keep the status of each statement clear, this
section is organized as follows: the growing-$k$ \emph{bracket} of \Cref{cor:bracket} is \emph{proved}; the \emph{sufficient
lemma} below that would close it is \emph{conditional} (we state it but do not prove it); the symmetric-route gap rests on a
proved estimate (\Cref{lem:bc}) but its conclusion is an obstruction to a proof method, not a lower bound; and the
phase-conveyor that blocks the asymmetric route is an \emph{informal} construction (\Cref{app:bc}). The \emph{exact}
growing-$k$ value remains \emph{open}. Nothing in this section is used by \Cref{thm:offline,thm:online,thm:separation}.

\paragraph{A sufficient lemma.}
Everything funnels through the rank statistic $\rank$ of \eqref{eq:rank}. Recall how the two kinds of online move act on it.
A range increase, which we call a ``raise,'' happens when the largest color load on a coordinate grows; it changes the rank
of that coordinate by $\Delta\rank=k-\maxmult$, where $\maxmult$ is the current multiplicity of the maximum. A within-level
broadening happens when a color that is one below the maximum is incremented, so the set of colors achieving the maximum
load grows by one; this enlarges the argmax set and changes the rank by $\Delta\rank=+1$ while leaving the range itself
fixed. Because the rank is built additively from these two kinds of increment, \eqref{eq:rank} translates a bound on the
rank directly into a bound on the discrepancy. Concretely, any online rule that keeps the rank uniformly bounded, in the
sense that $\max_{t,i}\rank_i(t)\le C(k\,\Psifn+\Hscale)$ for some constant $C$ over all times $t$ and coordinates $i$,
would imply $\ondisc_k=\Th(\max\{\Psifn,\loglog n/k\})$, which is exactly the conjectured order and closes the bracket. So a
single uniform rank bound is sufficient. The difficulty is that the rank accumulates from both raises and broadenings, and
these are controlled by very different mechanisms. The raises are handled by the aggregate level-raising cover bound, the
same device that drives the offline analysis, which limits how many coordinates can reach each successive range. The
broadenings are the missing piece: there is at present no comparably tight control on how often the argmax set of a
coordinate can grow, and it is precisely this within-level term that the two approaches below try, and fail, to bound.

\paragraph{Two obstructions.}
There are two natural ways to try to bound the broadenings, and each runs into a wall that we can describe exactly.

The first is the symmetric approach: track only the permutation-invariant pair $(\rangeop,\maxmult)$ and run a moment
hierarchy in those two coordinates. This is attractive because it is exactly the statistic the offline argument uses, and
the broadening cost is in fact controlled by an adaptive estimate in this language, namely \Cref{lem:bc}. The trouble is in
how that estimate composes with the raise cover. When the broadening bound of \Cref{lem:bc} is folded into the level-raising
recurrence, the two together produce a $k$-fold range recurrence of the form $\eta_{r+1}\lesssim(C\eta_r)^k$, where $\eta_r$
is the fraction of coordinates that have reached range $r$. Iterating this recurrence gives $\eta\le e^{-ck^{j}}$ after $j$
levels, so to make the failure probability beat a union bound over the $n$ coordinates one needs $k^{j}\gtrsim\log n$, that
is $j\gtrsim\loglog n/\log k$. The bracket, however, allows only $j$ up to $\loglog n/k$. The ratio of these two thresholds
is a factor of $k/\log k$, which is the $\log k$ gap: the symmetric recurrence overcounts by exactly a $\log k$ factor in the
exponent. This gap is not a slack in the bookkeeping that a sharper inequality could remove. A deterministic ``skip''
configuration, in which one coordinate per color is placed at the profile $(r,0,\dots,0)$, shows that the $k^{r}$ branching is
forced: no permutation-invariant rule that sees only $(\rangeop,\maxmult)$ can distinguish the colors well enough to avoid
it, because by symmetry it must treat all $k$ skip directions alike. So the symmetric route faces a concrete obstruction. We
stress that this is an obstruction to a particular proof method and not an online lower bound: it is evidence that the moment
hierarchy in $(\rangeop,\maxmult)$ cannot prove the conjecture, and says nothing about whether the conjecture is true.

The second approach is asymmetric: use a color-sensitive rule that breaks the symmetry the skip configuration exploits. Such
a rule recovers the correct branching, which is not the symmetric $k^{r}$ but the slower ``staggered Fibonacci'' recurrence
$F_k(m)=\sum_{j=1}^{k}F_k(m-j)=\Th(\phi_k^{m})$, whose growth rate $\phi_k\uparrow2$ as $k\to\infty$. This is the same
recurrence that governs the $\loglog n/k$ maximum load of \emph{Always-Go-Left} allocation \cite{BCSV2006}, where biasing
the tie-breaking toward a fixed direction replaces multiplicative branching by the gentler Fibonacci branching. With the
right branching in hand, the conjectured order would follow from a tail bound at the level of expectations, controlling the
expected number of coordinates reaching each phase. That expectation-level tail is, unfortunately, false. It is refuted by an
explicit construction we call the ``phase conveyor'' (\Cref{app:bc}): a deterministic schedule that recycles a small fixed
set of coordinates to manufacture consecutive phases using only linearly many arrivals, instead of the Fibonacci-many that a
typical instance would require. The conveyor inflates the expectation past what the desired bound allows. What saves the
conjecture is that the conveyor is fragile: the precise sequence of supports it needs has probability far below $n^{-A}$ for
every fixed $A$. So the conveyor refutes the expectation bound and the associated quenched factorization, and not the
high-probability statement, which is why the conjecture below survives even though the natural expectation-level proof of it
does not.

\begin{problem}\label{prob:growing}
Determine $\ondisc_k$ for growing $k$ ($k\le d$, $d,k=(\log n)^{o(1)}$). We conjecture
$\ondisc_k=\Th(\max\{\Psifn(\lam,\log k),\ \loglog n/k\})$. By \Cref{app:bc}, this follows from a truncated, cycle-aware
high-probability bound on the asymmetric rule's phase births.
\end{problem}
